
%\newcommand{\Coq}{\textsf{Coq}}
\newcommand{\javadoc}{\textsf{javadoc}}
\newcommand{\ocamldoc}{\textsf{ocamldoc}}
\newcommand{\coqdoc}{\textsf{coqdoc}}
\newcommand{\texmacs}{\TeX{}macs}
\newcommand{\monurl}[1]{#1}
%HEVEA\renewcommand{\monurl}[1]{\ahref{#1}{#1}}
%\newcommand{\lnot}{not} % Hevea handles these symbols nicely
%\newcommand{\lor}{or}
%\newcommand{\land}{\&}
%%% Beware : in a \texttt, -- is displayed as a unique - hence
%%% the following macro:
\newcommand{\mm}{\symbol{45}\symbol{45}}


\coqdoc\ is a documentation tool for the proof assistant
\Coq, similar to \javadoc\ or \ocamldoc. 
The task of \coqdoc\ is
\begin{enumerate}
\item to produce a nice \LaTeX\ and/or HTML document from the \Coq\ 
  sources, readable for a human and not only for the proof assistant;
\item to help the user navigating in his own (or third-party) sources.
\end{enumerate}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Principles}

Documentation is inserted into \Coq\ files as \emph{special comments}.  
Thus your files will compile as usual, whether you use \coqdoc\ or not.
\coqdoc\ presupposes that the given \Coq\ files are well-formed (at
least lexically).  Documentation starts with
\texttt{(**}, followed by a space, and ends with the pending \texttt{*)}. 
The documentation format is inspired
  by Todd~A.~Coram's \emph{Almost Free Text (AFT)} tool: it is mainly
ASCII text with some syntax-light controls, described below.
\coqdoc\ is robust: it shouldn't fail, whatever the input is. But
remember: ``garbage in, garbage out''.

\paragraph{\Coq\ material inside documentation.}
\Coq\ material is quoted between the
delimiters \texttt{[} and \texttt{]}. Square brackets may be nested,
the inner ones being understood as being part of the quoted code (thus
you can quote a term like \texttt{fun x => u} by writing
\texttt{[fun x => u]}). Inside quotations, the code is pretty-printed in
the same way as it is in code parts.

Pre-formatted vernacular is enclosed by \texttt{[[} and
\texttt{]]}. The former must be followed by a newline and the latter
must follow a newline.

\paragraph{Pretty-printing.}
\coqdoc\ uses different faces for identifiers and keywords.  
The pretty-printing of \Coq\ tokens (identifiers or symbols) can be
controlled using one of the following commands:
\begin{alltt}
(** printing \emph{token} %...\LaTeX...% #...HTML...# *)
\end{alltt}
or
\begin{alltt}
(** printing \emph{token} $...\LaTeX\ math...$ #...HTML...# *)
\end{alltt}
It gives the \LaTeX\ and HTML texts to be produced for the given \Coq\
token. One of the \LaTeX\ or HTML text may be omitted, causing the
default pretty-printing to be used for this token.

The printing for one token can be removed with
\begin{alltt}
(** remove printing \emph{token} *)
\end{alltt}

Initially, the pretty-printing table contains the following mapping:
\begin{center}
  \begin{tabular}{ll@{\qquad\qquad}ll@{\qquad\qquad}ll@{\qquad\qquad}}
    \verb!->!            & $\rightarrow$   &
    \verb!<-!            & $\leftarrow$    &
    \verb|*|             & $\times$        \\
    \verb|<=|            & $\le$           &
    \verb|>=|            & $\ge$           &
    \verb|=>|            & $\Rightarrow$   \\
    \verb|<>|            & $\not=$         &
    \verb|<->|           & $\leftrightarrow$ &
    \verb!|-!            & $\vdash$        \\
    \verb|\/|            & $\lor$          &
    \verb|/\|            & $\land$         &
    \verb|~|             & $\lnot$ 
  \end{tabular}
\end{center}
Any of these can be overwritten or suppressed using the
\texttt{printing} commands.

Important note: the recognition of tokens is done by a (ocaml)lex
automaton and thus applies the longest-match rule. For instance,
\verb!->~! is recognized as a single token, where \Coq\ sees two
tokens. It is the responsibility of the user to insert space between
tokens \emph{or} to give pretty-printing rules for the possible
combinations, e.g. 
\begin{verbatim}
(** printing ->~ %\ensuremath{\rightarrow\lnot}% *)
\end{verbatim}


\paragraph{Sections.}
Sections are introduced by 1 to 4 leading stars (i.e. at the beginning of the
line) followed by a space. One star is a section, two stars a sub-section, etc.
The section title is given on the remaining of the line.
Example:
\begin{verbatim}
    (** * Well-founded relations
  
        In this section, we introduce...  *)
\end{verbatim}


%TODO \paragraph{Fonts.}


\paragraph{Lists.}
List items are introduced by a leading dash.  \coqdoc\ uses whitespace
to determine the depth of a new list item and which text belongs in
which list items.  A list ends when a line of text starts at or before
the level of indenting of the list's dash.  A list item's dash must
always be the first non-space character on its line (so, in
particular, a list can not begin on the first line of a comment -
start it on the second line instead).

Example:
\begin{verbatim}
     We go by induction on [n]:
     - If [n] is 0...
     - If [n] is [S n'] we require...

       two paragraphs of reasoning, and two subcases:

       - In the first case...
       - In the second case...

     So the theorem holds.
\end{verbatim}

\paragraph{Rules.}
More than 4 leading dashes produce an horizontal rule.

\paragraph{Emphasis.}
Text can be italicized by placing it in underscores.  A non-identifier
character must precede the leading underscore and follow the trailing
underscore, so that uses of underscores in names aren't mistaken for
emphasis.  Usually, these are spaces or punctuation.

\begin{verbatim}
    This sentence contains some _emphasized text_.
\end{verbatim}

\paragraph{Escaping to \LaTeX\ and HTML.}
Pure \LaTeX\ or HTML material can be inserted using the following
escape sequences:
\begin{itemize}
\item \verb+$...LaTeX stuff...$+ inserts some \LaTeX\ material in math mode.
  Simply discarded in HTML output.

\item \verb+%...LaTeX stuff...%+ inserts some \LaTeX\ material.
  Simply discarded in HTML output.

\item \verb+#...HTML stuff...#+ inserts some HTML material. Simply
  discarded in \LaTeX\ output.
\end{itemize}

Note: to simply output the characters \verb+$+, \verb+%+ and \verb+#+
and escaping their escaping role, these characters must be doubled.

\paragraph{Verbatim.} 
Verbatim material is introduced by a leading \verb+<<+ and closed by
\verb+>>+ at the beginning of a line. Example:
\begin{verbatim}
Here is the corresponding caml code:
<<
  let rec fact n = 
    if n <= 1 then 1 else n * fact (n-1)
>>
\end{verbatim}


\paragraph{Hyperlinks.}
Hyperlinks can be inserted into the HTML output, so that any
identifier is linked to the place of its definition.

\texttt{coqc \emph{file}.v} automatically dumps localization information
in \texttt{\emph{file}.glob} or appends it to a file specified using option
\texttt{\mm{}dump-glob \emph{file}}. Take care of erasing this global file, if
any, when starting the whole compilation process.

Then invoke \texttt{coqdoc} or \texttt{coqdoc \mm{}glob-from \emph{file}} to tell
\coqdoc\ to look for name resolutions into the file \texttt{\emph{file}}
(it will look in \texttt{\emph{file}.glob} by default).

Identifiers from the \Coq\ standard library are linked to the \Coq\
web site at \url{http://coq.inria.fr/library/}. This behavior can be
changed using command line options \texttt{\mm{}no-externals} and
\texttt{\mm{}coqlib}; see below.


\paragraph{Hiding / Showing parts of the source.}
Some parts of the source can be hidden using command line options
\texttt{-g} and \texttt{-l} (see below), or using such comments:
\begin{alltt}
(* begin hide *)
\emph{some Coq material}
(* end hide *)
\end{alltt}
Conversely, some parts of the source which would be hidden can be
shown using such comments: 
\begin{alltt}
(* begin show *)
\emph{some Coq material}
(* end show *)
\end{alltt}
The latter cannot be used around some inner parts of a proof, but can
be used around a whole proof.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Usage}

\coqdoc\ is invoked on a shell command line as follows:
\begin{displaymath}
  \texttt{coqdoc }<\textit{options and files}>
\end{displaymath}
Any command line argument which is not an option is considered to be a
file (even if it starts with a \verb!-!). \Coq\ files are identified
by the suffixes \verb!.v! and \verb!.g! and \LaTeX\ files by the
suffix \verb!.tex!. 

\begin{description}
\item[HTML output] ~\par
  This is the default output.
  One HTML file is created for each \Coq\ file given on the command line,
  together with a file \texttt{index.html} (unless option
  \texttt{-no-index} is passed). The HTML pages use a style sheet
  named \texttt{style.css}. Such a file is distributed with \coqdoc.

\item[\LaTeX\ output] ~\par
  A single \LaTeX\ file is created, on standard output. It can be
  redirected to a file with option \texttt{-o}. 
  The order of files on the command line is kept in the final
  document. \LaTeX\ files given on the command line are copied `as is'
  in the final document .
  DVI and PostScript can be produced directly with the options
  \texttt{-dvi} and \texttt{-ps} respectively.

\item[\texmacs\ output] ~\par
  To translate the input files to \texmacs\ format, to be used by
  the \texmacs\ Coq interface 
  (see \url{http://www-sop.inria.fr/lemme/Philippe.Audebaud/tmcoq/}).
\end{description}


\subsubsection*{Command line options}


\paragraph{Overall options}

\begin{description}

\item[\texttt{\mm{}html}] ~\par
  
  Select a HTML output.

\item[\texttt{\mm{}latex}] ~\par
  
  Select a \LaTeX\ output.

\item[\texttt{\mm{}dvi}] ~\par
  
  Select a DVI output.

\item[\texttt{\mm{}ps}] ~\par
  
  Select a PostScript output.

\item[\texttt{\mm{}texmacs}] ~\par
  
  Select a \texmacs\ output.

\item[\texttt{--stdout}] ~\par

  Write output to stdout.

\item[\texttt{-o }\textit{file}, \texttt{\mm{}output }\textit{file}] ~\par
  
  Redirect the output into the file `\textit{file}' (meaningless with
  \texttt{-html}).

\item[\texttt{-d }\textit{dir}, \texttt{\mm{}directory }\textit{dir}] ~\par

  Output files into directory `\textit{dir}' instead of current
  directory (option \texttt{-d} does not change the filename specified
  with option \texttt{-o}, if any).

\item[\texttt{\mm{}body-only}] ~\par

  Suppress the header and trailer of the final document. Thus, you can
  insert the resulting document into a larger one.

\item[\texttt{-p} \textit{string}, \texttt{\mm{}preamble} \textit{string}]~\par

  Insert some material in the \LaTeX\ preamble, right before
  \verb!\begin{document}! (meaningless with \texttt{-html}).

\item[\texttt{\mm{}vernac-file }\textit{file},
      \texttt{\mm{}tex-file }\textit{file}] ~\par
      
      Considers the file `\textit{file}' respectively as a \verb!.v!
      (or \verb!.g!) file or a \verb!.tex! file.

\item[\texttt{\mm{}files-from }\textit{file}] ~\par

  Read file names to process in file `\textit{file}' as if they were
  given on the command line. Useful for program sources split up into
  several directories.
  
\item[\texttt{-q}, \texttt{\mm{}quiet}] ~\par

  Be quiet. Do not print anything except errors.

\item[\texttt{-h}, \texttt{\mm{}help}] ~\par

  Give a short summary of the options and exit.

\item[\texttt{-v}, \texttt{\mm{}version}] ~\par

  Print the version and exit.

\end{description}

\paragraph{Index options}

Default behavior is to build an index, for the HTML output only, into
\texttt{index.html}.

\begin{description}

\item[\texttt{\mm{}no-index}] ~\par
  
  Do not output the index.

\item[\texttt{\mm{}multi-index}] ~\par
  
  Generate one page for each category and each letter in the index,
  together with a top page \texttt{index.html}.

\item[\texttt{\mm{}index }\textit{string}] ~\par

  Make the filename of the index \textit{string} instead of ``index''.
  Useful since ``index.html'' is special.

\end{description}

\paragraph{Table of contents option}

\begin{description}

\item[\texttt{-toc}, \texttt{\mm{}table-of-contents}] ~\par

  Insert a table of contents.
  For a \LaTeX\ output, it inserts a \verb!\tableofcontents! at the
  beginning of the document. For a HTML output, it builds a table of
  contents into \texttt{toc.html}.

\item[\texttt{\mm{}toc-depth }\textit{int}] ~\par

  Only include headers up to depth \textit{int} in the table of
  contents.

\end{description}

\paragraph{Hyperlinks options}
\begin{description}

\item[\texttt{\mm{}glob-from }\textit{file}] ~\par
  
  Make references using \Coq\ globalizations from file \textit{file}. 
  (Such globalizations are obtained with \Coq\ option \texttt{-dump-glob}).

\item[\texttt{\mm{}no-externals}] ~\par
  
  Do not insert links to the \Coq\ standard library.

\item[\texttt{\mm{}external }\textit{url}~\textit{coqdir}] ~\par
  
  Use given URL for linking references whose name starts with prefix
  \textit{coqdir}.

\item[\texttt{\mm{}coqlib }\textit{url}] ~\par

  Set base URL for the \Coq\ standard library (default is 
  \url{http://coq.inria.fr/library/}). This is equivalent to
  \texttt{\mm{}external }\textit{url}~\texttt{Coq}.

\item[\texttt{-R }\textit{dir }\textit{coqdir}] ~\par

  Map physical directory \textit{dir} to \Coq\ logical directory
  \textit{coqdir} (similarly to \Coq\ option \texttt{-R}).

  Note: option \texttt{-R} only has effect on the files
  \emph{following} it on the command line, so you will probably need
  to put this option first.

\end{description}

\paragraph{Title options}
\begin{description}
\item[\texttt{-s }, \texttt{\mm{}short}] ~\par
  
  Do not insert titles for the files. The default behavior is to
  insert a title like ``Library Foo'' for each file.

\item[\texttt{\mm{}lib-name }\textit{string}] ~\par

  Print ``\textit{string} Foo'' instead of ``Library Foo'' in titles.
  For example ``Chapter'' and ``Module'' are reasonable choices.

\item[\texttt{\mm{}no-lib-name}] ~\par

  Print just ``Foo'' instead of ``Library Foo'' in titles.

\item[\texttt{\mm{}lib-subtitles}] ~\par

  Look for library subtitles.  When enabled, the beginning of each
  file is checked for a comment of the form:
\begin{alltt}
(** * ModuleName : text *)
\end{alltt}
  where \texttt{ModuleName} must be the name of the file.  If it is
  present, the \texttt{text} is used as a subtitle for the module in
  appropriate places.

\item[\texttt{-t }\textit{string}, 
      \texttt{\mm{}title }\textit{string}] ~\par
  
  Set the document title.      

\end{description}

\paragraph{Contents options}
\begin{description}

\item[\texttt{-g}, \texttt{\mm{}gallina}] ~\par

  Do not print proofs.

\item[\texttt{-l}, \texttt{\mm{}light}] ~\par
  
  Light mode. Suppress proofs (as with \texttt{-g}) and the following commands:
  \begin{itemize}
  \item {}[\texttt{Recursive}] \texttt{Tactic Definition}
  \item \texttt{Hint / Hints} 
  \item \texttt{Require}
  \item \texttt{Transparent / Opaque}
  \item \texttt{Implicit Argument / Implicits}
  \item \texttt{Section / Variable / Hypothesis / End}
  \end{itemize}

\end{description}
The behavior of options \texttt{-g} and \texttt{-l} can be locally
overridden using the \texttt{(* begin show *)} \dots\ \texttt{(* end
  show *)} environment (see above).

There are a few options to drive the parsing of comments:
\begin{description}
\item[\texttt{\mm{}parse-comments}] ~\par

  Parses regular comments delimited by \texttt{(*} and \texttt{*)} as 
  well. They are typeset inline.

\item[\texttt{\mm{}plain-comments}] ~\par

  Do not interpret comments, simply copy them as plain-text.

\item[\texttt{\mm{}interpolate}] ~\par

  Use the globalization information to typeset identifiers appearing in
  \Coq{} escapings inside comments.
\end{description}


\paragraph{Language options}

Default behavior is to assume ASCII 7 bits input files.

\begin{description}

\item[\texttt{-latin1}, \texttt{\mm{}latin1}] ~\par

  Select ISO-8859-1 input files. It is equivalent to
  \texttt{--inputenc latin1 --charset iso-8859-1}.

\item[\texttt{-utf8}, \texttt{\mm{}utf8}] ~\par

  Select UTF-8 (Unicode) input files. It is equivalent to
  \texttt{--inputenc utf8 --charset utf-8}.
  \LaTeX\ UTF-8 support can be found at
 \url{http://www.ctan.org/pkg/unicode}.

\item[\texttt{\mm{}inputenc} \textit{string}] ~\par

  Give a \LaTeX\ input encoding, as an option to \LaTeX\ package
  \texttt{inputenc}. 

\item[\texttt{\mm{}charset} \textit{string}] ~\par

  Specify the HTML character set, to be inserted in the HTML header.

\end{description}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection[The coqdoc \LaTeX{} style file]{The coqdoc \LaTeX{} style file\label{section:coqdoc.sty}}

In case you choose to produce a document without the default \LaTeX{}
preamble (by using option \verb|--no-preamble|), then you must insert
into your own preamble the command
\begin{quote}
  \verb|\usepackage{coqdoc}|
\end{quote}

The package optionally takes the argument \verb|[color]| to typeset
identifiers with colors (this requires the \verb|xcolor| package).

Then you may alter the rendering of the document by
redefining some macros:
\begin{description}

\item[\texttt{coqdockw}, \texttt{coqdocid}, \ldots] ~ 
  
  The one-argument macros for typesetting keywords and identifiers.
  Defaults are sans-serif for keywords and italic for identifiers.

  For example, if you would like a slanted font for keywords, you
  may insert  
\begin{verbatim}
     \renewcommand{\coqdockw}[1]{\textsl{#1}}
\end{verbatim}
  anywhere between \verb|\usepackage{coqdoc}| and
  \verb|\begin{document}|. 

\item[\texttt{coqdocmodule}] ~ 
  
  One-argument macro for typesetting the title of a \verb|.v| file.
  Default is
\begin{verbatim}
\newcommand{\coqdocmodule}[1]{\section*{Module #1}}
\end{verbatim}
  and you may redefine it using \verb|\renewcommand|.

\end{description}


