\chapter[Calculus of Inductive Constructions]{Calculus of Inductive Constructions
\label{Cic}
\index{Cic@\textsc{CIC}}
\index{Calculus of Inductive Constructions}}

The underlying formal language of {\Coq} is a {\em Calculus of
Inductive Constructions} (\CIC) whose inference rules are presented in
this chapter. The history of this formalism as well as pointers to related work
are provided in a separate chapter; see {\em Credits}.

\section[The terms]{The terms\label{Terms}}

The expressions of the {\CIC} are {\em terms} and all terms have a {\em type}.
There are types for functions (or
programs), there are atomic types (especially datatypes)... but also
types for proofs and types for the types themselves.
Especially, any object handled in the formalism must belong to a
type.  For instance, universal quantification is relative to a type and
takes the form {\it ``for all x
of type T, P''}. The expression {\it ``x of type T''} is
written {\it ``x:T''}. Informally, {\it ``x:T''} can be thought as
{\it ``x belongs to T''}.

The types of types are {\em sorts}. Types and sorts are themselves
terms so that terms, types and sorts are all components of a common
syntactic language of terms which is described in
Section~\ref{cic:terms} but, first, we describe sorts.

\subsection[Sorts]{Sorts\label{Sorts}
\index{Sorts}}
All sorts have a type and there is an infinite well-founded
typing hierarchy of sorts whose base sorts are {\Prop} and {\Set}.

The sort {\Prop} intends to be the type of logical propositions. If
$M$ is a logical proposition then it denotes the class of terms
representing proofs of $M$. An object $m$ belonging to $M$ witnesses
the fact that $M$ is provable. An object of type {\Prop} is called a
proposition.

The sort {\Set} intends to be the type of small sets. This includes data
types such as booleans and naturals, but also products, subsets, and
function types over these data types.

{\Prop} and {\Set} themselves can be manipulated as ordinary
terms. Consequently they also have a type. Because assuming simply
that {\Set} has type {\Set} leads to an inconsistent theory~\cite{Coq86}, the
language of {\CIC} has infinitely many sorts. There are, in addition
to {\Set} and {\Prop} a hierarchy of universes {\Type$(i)$} for any
integer $i$.

Like {\Set}, all of the sorts {\Type$(i)$} contain small sets such as
booleans, natural numbers, as well as products, subsets and function
types over small sets. But, unlike {\Set}, they also contain large
sets, namely the sorts {\Set} and {\Type$(j)$} for $j<i$, and all
products, subsets and function types over these sorts.

Formally, we call {\Sort} the set of sorts which is defined by:
\index{Type@{\Type}}%
\index{Prop@{\Prop}}%
\index{Set@{\Set}}%
\[\Sort \equiv \{\Prop,\Set,\Type(i)\;|\; i \in \NN\} \]
Their properties, such as:
{\Prop:\Type$(1)$}, {\Set:\Type$(1)$}, and {\Type$(i)$:\Type$(i+1)$},
are defined in Section~\ref{subtyping-rules}.

The user does not have to mention explicitly the index $i$ when referring to
the universe \Type$(i)$. One only writes \Type. The
system itself generates for each instance of \Type\ a new
index for the universe and checks that the constraints between these
indexes can be solved. From the user point of view we consequently
have {\Type}:{\Type}.
We shall make precise in the typing rules the constraints between the
indexes. 

\paragraph{Implementation issues}
In practice, the {\Type} hierarchy is implemented using
{\em algebraic universes}\index{algebraic universe}.
An algebraic universe $u$ is either a variable (a qualified
identifier with a number) or a successor of an algebraic universe (an
expression $u+1$), or an upper bound of algebraic universes (an
expression $max(u_1,...,u_n)$), or the base universe (the expression
$0$) which corresponds, in the arity of template polymorphic inductive
types (see Section \ref{Template-polymorphism}),
to the predicative sort {\Set}. A graph of constraints between
the universe variables is maintained globally. To ensure the existence
of a mapping of the universes to the positive integers, the graph of
constraints must remain acyclic.  Typing expressions that violate the
acyclicity of the graph of constraints results in a \errindex{Universe
inconsistency} error (see also Section~\ref{PrintingUniverses}).

%% HH: This looks to me more like source of confusion than helpful

%% \subsection{Constants}

%% Constants refers to
%% objects in the global environment. These constants may denote previously
%% defined objects, but also objects related to inductive definitions
%% (either the type itself or one of its constructors or destructors).

%% \medskip\noindent {\bf Remark. } In other presentations of \CIC, 
%% the inductive objects are not seen as
%% external declarations but as first-class terms. Usually the
%% definitions are also completely ignored.  This is a nice theoretical
%% point of view but not so practical. An inductive definition is
%% specified by a possibly huge set of declarations, clearly we want to
%% share this specification among the various inductive objects and not
%% to duplicate it. So the specification should exist somewhere and the
%% various objects should refer to it.  We choose one more level of
%% indirection where the objects are just represented as constants and
%% the environment gives the information on the kind of object the
%% constant refers to.

%% \medskip
%% Our inductive objects will be manipulated as constants declared in the
%% environment. This roughly corresponds to the way they are actually
%% implemented in the \Coq\ system. It is simple to map this presentation
%% in a theory where inductive objects are represented by terms.

\subsection{Terms}
\label{cic:terms}

Terms are built from sorts, variables, constants,
%constructors, inductive types,
abstractions, applications, local definitions,
%case analysis, fixpoints, cofixpoints
and products.
From a syntactic point of view, types cannot be distinguished from terms,
except that they cannot start by an abstraction or a constructor.
More precisely the language of the {\em Calculus of Inductive
  Constructions} is built from the following rules.
%
\begin{enumerate}
\item the sorts {\Set}, {\Prop}, ${\Type(i)}$ are terms.
\item variables, hereafter ranged over by letters $x$, $y$, etc., are terms
\item constants, hereafter ranged over by letters $c$, $d$, etc.,  are terms.
%\item constructors, hereafter ranged over by letter $C$, are terms.
%\item inductive types, hereafter ranged over by letter $I$, are terms.
\item\index{products} if $x$ is a variable and $T$, $U$ are terms then $\forall~x:T,U$
  ($\kw{forall}~x:T,~U$ in \Coq{} concrete syntax) is a term. If $x$
  occurs in $U$, $\forall~x:T,U$ reads as {\it ``for all x of type T,
    U''}. As $U$ depends on $x$, one says that $\forall~x:T,U$ is a
  {\em dependent product}. If $x$ does not occur in $U$ then
  $\forall~x:T,U$ reads as {\it ``if T then U''}. A {\em non dependent
  product} can be written: $T \ra U$.
\item if $x$ is a variable and $T$, $u$ are terms then $\lb x:T \mto u$
  ($\kw{fun}~x:T~ {\tt =>}~ u$ in \Coq{} concrete syntax) is a term. This is a
  notation for the $\lambda$-abstraction of
  $\lambda$-calculus\index{lambda-calculus@$\lambda$-calculus}
  \cite{Bar81}. The term $\lb x:T \mto u$ is a function which maps
  elements of $T$ to the expression $u$.
\item if $t$ and $u$ are terms then $(t\ u)$ is a term  
 ($t~u$ in \Coq{} concrete syntax).  The term $(t\ 
  u)$ reads as {\it ``t applied to u''}.
\item if $x$ is a variable, and $t$, $T$ and $u$ are terms then
  $\kw{let}~x:=t:T~\kw{in}~u$ is a
  term which denotes the term $u$ where the variable $x$ is locally
  bound to $t$ of type $T$. This stands for the common ``let-in''
  construction of functional programs such as ML or Scheme.
%\item case ...
%\item fixpoint ...
%\item cofixpoint ...
\end{enumerate}

\paragraph{Free variables.}
The notion of free variables is defined as usual.  In the expressions
$\lb x:T\mto U$ and $\forall x:T, U$ the occurrences of $x$ in $U$
are bound.

\paragraph[Substitution.]{Substitution.\index{Substitution}}
The notion of substituting a term $t$ to free occurrences of a
variable $x$ in a term $u$ is defined as usual. The resulting term
is written $\subst{u}{x}{t}$.

\paragraph[The logical vs programming readings.]{The logical vs programming readings.}

The constructions of the {\CIC} can be used to express both logical
and programming notions, accordingly to the Curry-Howard
correspondence between proofs and programs, and between propositions
and types~\cite{Cur58,How80,Bru72}.

For instance, let us assume that \nat\ is the type of natural numbers
with zero element written $0$ and that ${\tt True}$ is the always true
proposition.  Then $\ra$ is used both to denote $\nat\ra\nat$ which is
the type of functions from \nat\ to \nat, to denote ${\tt True}\ra{\tt
  True}$ which is an implicative proposition, to denote $\nat \ra
\Prop$ which is the type of unary predicates over the natural numbers,
etc.

Let us assume that ${\tt mult}$ is a function of type $\nat\ra\nat\ra
\nat$ and ${\tt eqnat}$ a predicate of type $\nat\ra\nat\ra \Prop$.
The $\lambda$-abstraction can serve to build ``ordinary'' functions as
in $\lambda x:\nat.({\tt mult}~x~x)$ (i.e. $\kw{fun}~x:\nat ~{\tt =>}~
{\tt mult} ~x~x$ in {\Coq} notation) but may build also predicates
over the natural numbers. For instance $\lambda x:\nat.({\tt eqnat}~
x~0)$ (i.e. $\kw{fun}~x:\nat ~{\tt =>}~ {\tt eqnat}~ x~0$ in {\Coq}
notation) will represent the predicate of one variable $x$ which
asserts the equality of $x$ with $0$. This predicate has type $\nat
\ra \Prop$ and it can be applied to any expression of type ${\nat}$,
say $t$, to give an object $P~t$ of type \Prop, namely a proposition.

Furthermore $\kw{forall}~x:\nat,\,P\;x$ will represent the type of
functions which associate to each natural number $n$ an object of type
$(P~n)$ and consequently represent the type of proofs of the formula
``$\forall x.\,P(x)$''.

\section[Typing rules]{Typing rules\label{Typed-terms}}

As objects of type theory, terms are subjected to {\em type
discipline}. The well typing of a term depends on
a global environment and a local context.

\paragraph{Local context.\index{Local context}}
A {\em local context} is an ordered list of
{\em local declarations\index{declaration!local}} of names which we call {\em variables\index{variable}}.
The declaration of some variable $x$ is
either a {\em local assumption\index{assumption!local}}, written $x:T$ ($T$ is a type) or a {\em local definition\index{definition!local}},
written $x:=t:T$.  We use brackets to write local contexts. A
typical example is $[x:T;y:=u:U;z:V]$.  Notice that the variables
declared in a local context must be distinct. If $\Gamma$ declares some $x$,
we write $x \in \Gamma$. By writing $(x:T) \in \Gamma$ we mean that
either $x:T$ is an assumption in $\Gamma$ or that there exists some $t$ such
that $x:=t:T$ is a definition in $\Gamma$. If $\Gamma$ defines some
$x:=t:T$, we also write $(x:=t:T) \in \Gamma$.
For the rest of the chapter, the $\Gamma::(y:T)$ denotes the local context
$\Gamma$ enriched with the local assumption $y:T$.
Similarly, $\Gamma::(y:=t:T)$ denotes the local context
$\Gamma$ enriched with the local definition $(y:=t:T)$.
The notation $[]$ denotes the empty local context.
By $\Gamma_1; \Gamma_2$ we mean concatenation of the local context $\Gamma_1$
and the local context $\Gamma_2$.

% Does not seem to be used further...
% Si dans l'explication WF(E)[Gamma] concernant les constantes
% definies ds un contexte

%We define the inclusion of two local contexts $\Gamma$ and $\Delta$ (written
%as $\Gamma \subset \Delta$) as the property, for all variable $x$,
%type $T$ and term $t$, if $(x:T) \in \Gamma$ then $(x:T) \in \Delta$
%and if $(x:=t:T) \in \Gamma$ then $(x:=t:T) \in \Delta$.
%We write
% $|\Delta|$ for the length of the context $\Delta$, that is for the number
% of declarations (assumptions or definitions) in $\Delta$.

\paragraph[Global environment.]{Global environment.\index{Global environment}}
%Because we are manipulating global declarations (global constants and global
%assumptions), we also need to consider a global environment $E$.

A {\em global environment} is an ordered list of {\em global declarations\index{declaration!global}}.
Global declarations are either {\em global assumptions\index{assumption!global}} or {\em global
definitions\index{definition!global}}, but also declarations of inductive objects. Inductive objects themselves declare both inductive or coinductive types and constructors
(see Section~\ref{Cic-inductive-definitions}).

A {\em global assumption} will be represented in the global environment as
$(c:T)$ which assumes the name $c$ to be of some type $T$.
A {\em global definition} will
be represented in the global environment as $c:=t:T$ which defines
the name $c$ to have value $t$ and type $T$.
We shall call such names {\em constants}.
For the rest of the chapter, the $E;c:T$ denotes the global environment
$E$ enriched with the global assumption $c:T$.
Similarly, $E;c:=t:T$ denotes the global environment
$E$ enriched with the global definition $(c:=t:T)$.

The rules for inductive definitions (see Section
\ref{Cic-inductive-definitions}) have to be considered as assumption
rules to which the following definitions apply: if the name $c$ is
declared in $E$, we write $c \in E$ and if $c:T$ or $c:=t:T$ is
declared in $E$, we write $(c : T) \in E$.

\paragraph[Typing rules.]{Typing rules.\label{Typing-rules}\index{Typing rules}}
In the following, we define simultaneously two
judgments.  The first one \WTEG{t}{T} means the term $t$ is well-typed
and has type $T$ in the global environment $E$ and local context $\Gamma$.  The
second judgment \WFE{\Gamma} means that the global environment $E$ is
well-formed and the local context $\Gamma$ is a valid local context in this
global environment.
% HH: This looks to me complicated. I think it would be better to talk
% about ``discharge'' as a transformation of global environments,
% rather than as keeping a local context next to global constants.
%
%%  It also means a third property which makes sure that any
%%constant in $E$ was defined in an environment which is included in
%%$\Gamma$
%%\footnote{This requirement could be relaxed if we instead introduced
%%  an explicit mechanism for instantiating constants. At the external
%%  level, the Coq engine works accordingly to this view that all the
%%  definitions in the environment were built in a local sub-context of the
%%  current local context.}.

A term $t$ is well typed in a global environment $E$ iff there exists a
local context $\Gamma$ and a term $T$ such that the judgment \WTEG{t}{T} can
be derived from the following rules.
\begin{description}
\item[W-Empty] \inference{\WF{[]}{}}
\item[W-Local-Assum]  % Ce n'est pas vrai : x peut apparaitre plusieurs fois dans Gamma
\inference{\frac{\WTEG{T}{s}~~~~s \in \Sort~~~~x \not\in \Gamma % \cup E
      }{\WFE{\Gamma::(x:T)}}}
\item[W-Local-Def]
\inference{\frac{\WTEG{t}{T}~~~~x \not\in \Gamma % \cup E
     }{\WFE{\Gamma::(x:=t:T)}}}
\item[W-Global-Assum] \inference{\frac{\WTE{}{T}{s}~~~~s \in \Sort~~~~c \notin E}
                      {\WF{E;c:T}{}}}
\item[W-Global-Def] \inference{\frac{\WTE{}{t}{T}~~~c \notin E}
                      {\WF{E;c:=t:T}{}}}
\item[Ax-Prop] \index{Typing rules!Ax-Prop}
\inference{\frac{\WFE{\Gamma}}{\WTEG{\Prop}{\Type(1)}}}
\item[Ax-Set] \index{Typing rules!Ax-Set}
\inference{\frac{\WFE{\Gamma}}{\WTEG{\Set}{\Type(1)}}}
\item[Ax-Type] \index{Typing rules!Ax-Type}
\inference{\frac{\WFE{\Gamma}}{\WTEG{\Type(i)}{\Type(i+1)}}}
\item[Var]\index{Typing rules!Var}
 \inference{\frac{ \WFE{\Gamma}~~~~~(x:T) \in \Gamma~~\mbox{or}~~(x:=t:T) \in \Gamma~\mbox{for some $t$}}{\WTEG{x}{T}}}
\item[Const]  \index{Typing rules!Const}
\inference{\frac{\WFE{\Gamma}~~~~(c:T) \in E~~\mbox{or}~~(c:=t:T) \in E~\mbox{for some $t$} }{\WTEG{c}{T}}}
\item[Prod-Prop]  \index{Typing rules!Prod-Prop}
\inference{\frac{\WTEG{T}{s}~~~~s \in \Sort~~~
    \WTE{\Gamma::(x:T)}{U}{\Prop}}
      { \WTEG{\forall~x:T,U}{\Prop}}} 
\item[Prod-Set]  \index{Typing rules!Prod-Set}
\inference{\frac{\WTEG{T}{s}~~~~s \in\{\Prop, \Set\}~~~~~~
    \WTE{\Gamma::(x:T)}{U}{\Set}}
      { \WTEG{\forall~x:T,U}{\Set}}} 
\item[Prod-Type]  \index{Typing rules!Prod-Type}
\inference{\frac{\WTEG{T}{\Type(i)}~~~~
    \WTE{\Gamma::(x:T)}{U}{\Type(i)}}
    {\WTEG{\forall~x:T,U}{\Type(i)}}}
\item[Lam]\index{Typing rules!Lam} 
\inference{\frac{\WTEG{\forall~x:T,U}{s}~~~~ \WTE{\Gamma::(x:T)}{t}{U}}
        {\WTEG{\lb x:T\mto t}{\forall x:T, U}}}
\item[App]\index{Typing rules!App}
 \inference{\frac{\WTEG{t}{\forall~x:U,T}~~~~\WTEG{u}{U}}
                 {\WTEG{(t\ u)}{\subst{T}{x}{u}}}}
\item[Let]\index{Typing rules!Let} 
\inference{\frac{\WTEG{t}{T}~~~~ \WTE{\Gamma::(x:=t:T)}{u}{U}}
        {\WTEG{\letin{x}{t:T}{u}}{\subst{U}{x}{t}}}}
\end{description}

\Rem Prod$_1$ and Prod$_2$ typing-rules make sense if we consider the semantic
difference between {\Prop} and {\Set}:
\begin{itemize}
  \item All values of a type that has a sort {\Set} are extractable.
  \item No values of a type that has a sort {\Prop} are extractable.
\end{itemize}

\Rem We may have $\kw{let}~x:=t:T~\kw{in}~u$
well-typed without having $((\lb x:T\mto u)~t)$ well-typed (where
$T$ is a type of $t$). This is because the value $t$ associated to $x$
may be used in a conversion rule (see Section~\ref{conv-rules}).

\section[Conversion rules]{Conversion rules\index{Conversion rules}
\label{conv-rules}}

In \CIC, there is an internal reduction mechanism. In particular, it
can decide if two programs are {\em intentionally} equal (one
says {\em convertible}). Convertibility is described in this section.

\paragraph[$\beta$-reduction.]{$\beta$-reduction.\label{beta}\index{beta-reduction@$\beta$-reduction}}

We want to be able to identify some terms as we can identify the
application of a function to a given argument with its result. For
instance the identity function over a given type $T$ can be written
$\lb x:T\mto x$. In any global environment $E$ and local context $\Gamma$, we want to identify any object $a$ (of type $T$) with the
application $((\lb x:T\mto x)~a)$. We define for this a {\em reduction} (or a
{\em conversion}) rule we call $\beta$:
\[ \WTEGRED{((\lb x:T\mto
  t)~u)}{\triangleright_{\beta}}{\subst{t}{x}{u}} \] 
We say that $\subst{t}{x}{u}$ is the {\em $\beta$-contraction} of
$((\lb x:T\mto t)~u)$ and, conversely, that $((\lb x:T\mto t)~u)$
is the {\em $\beta$-expansion} of $\subst{t}{x}{u}$.

According to $\beta$-reduction, terms of the {\em Calculus of
  Inductive Constructions} enjoy some fundamental properties such as
confluence, strong normalization, subject reduction. These results are
theoretically of great importance but we will not detail them here and
refer the interested reader to \cite{Coq85}.

\paragraph[$\iota$-reduction.]{$\iota$-reduction.\label{iota}\index{iota-reduction@$\iota$-reduction}}
A specific conversion rule is associated to the inductive objects in
the global environment.  We shall give later on (see Section~\ref{iotared}) the
precise rules but it just says that a destructor applied to an object
built from a constructor behaves as expected.  This reduction is
called $\iota$-reduction and is more precisely studied in
\cite{Moh93,Wer94}.


\paragraph[$\delta$-reduction.]{$\delta$-reduction.\label{delta}\index{delta-reduction@$\delta$-reduction}}

We may have variables defined in local contexts or constants defined in the global
environment. It is legal to identify such a reference with its value,
that is to expand (or unfold) it into its value. This
reduction is called $\delta$-reduction and shows as follows.

$$\WTEGRED{x}{\triangleright_{\delta}}{t}~~~~~\mbox{if $(x:=t:T) \in \Gamma$}~~~~~~~~~\WTEGRED{c}{\triangleright_{\delta}}{t}~~~~~\mbox{if $(c:=t:T) \in E$}$$


\paragraph[$\zeta$-reduction.]{$\zeta$-reduction.\label{zeta}\index{zeta-reduction@$\zeta$-reduction}}

{\Coq} allows also to remove local definitions occurring in terms by
replacing the defined variable by its value. The declaration being
destroyed, this reduction differs from $\delta$-reduction. It is
called $\zeta$-reduction and shows as follows.

$$\WTEGRED{\kw{let}~x:=u~\kw{in}~t}{\triangleright_{\zeta}}{\subst{t}{x}{u}}$$

\paragraph{$\eta$-expansion.%
\label{eta}%
\index{eta-expansion@$\eta$-expansion}%
%\index{eta-reduction@$\eta$-reduction}
}%
Another important concept is $\eta$-expansion. It is legal to identify any
term $t$ of functional type $\forall x:T, U$ with its so-called
$\eta$-expansion $\lb x:T\mto (t\ x)$ for $x$ an arbitrary variable
name fresh in $t$.

\Rem We deliberately do not define $\eta$-reduction:
\begin{latexonly}%
  $$\lb x:T\mto (t\ x)\not\triangleright_\eta\hskip.3em t$$
\end{latexonly}%
\begin{htmlonly}
  $$\lb x:T\mto (t\ x)~\not\triangleright_\eta~t$$
\end{htmlonly}
This is because, in general, the type of $t$ need not to be convertible to the type of $\lb x:T\mto (t\ x)$.
E.g., if we take $f$ such that:
\begin{latexonly}%
  $$f\hskip.5em:\hskip.5em\forall x:Type(2),Type(1)$$
\end{latexonly}%
\begin{htmlonly}
  $$f~:~\forall x:Type(2),Type(1)$$
\end{htmlonly}
then
\begin{latexonly}%
  $$\lb x:Type(1),(f\, x)\hskip.5em:\hskip.5em\forall x:Type(1),Type(1)$$
\end{latexonly}%
\begin{htmlonly}
  $$\lb x:Type(1),(f\, x)~:~\forall x:Type(1),Type(1)$$
\end{htmlonly}
We could not allow
\begin{latexonly}%
  $$\lb x:Type(1),(f\,x)\hskip.4em\not\triangleright_\eta\hskip.6em f$$
\end{latexonly}%
\begin{htmlonly}
  $$\lb x:Type(1),(f\,x)~\not\triangleright_\eta~f$$
\end{htmlonly}
because the type of the reduced term $\forall x:Type(2),Type(1)$
would not be convertible to the type of the original term $\forall x:Type(1),Type(1)$.

\paragraph[Convertibility.]{Convertibility.\label{convertibility}
\index{beta-reduction@$\beta$-reduction}\index{iota-reduction@$\iota$-reduction}\index{delta-reduction@$\delta$-reduction}\index{zeta-reduction@$\zeta$-reduction}}

Let us write $\WTEGRED{t}{\triangleright}{u}$ for the contextual closure of the relation $t$ reduces to $u$ in the global environment $E$ and local context $\Gamma$ with one of the previous reduction $\beta$, $\iota$, $\delta$ or $\zeta$.

We say that two terms $t_1$ and $t_2$ are {\em
  $\beta\iota\delta\zeta\eta$-convertible}, or simply {\em
  convertible}, or {\em equivalent}, in the global environment $E$ and
local context $\Gamma$ iff there exist terms $u_1$ and $u_2$ such that
$\WTEGRED{t_1}{\triangleright \ldots \triangleright}{u_1}$ and
$\WTEGRED{t_2}{\triangleright \ldots \triangleright}{u_2}$ and either
$u_1$ and $u_2$ are identical, or they are convertible up to
$\eta$-expansion, i.e. $u_1$ is $\lb x:T\mto u'_1$ and $u_2\,x$ is
recursively convertible to $u'_1$, or, symmetrically, $u_2$ is $\lb
x:T\mto u'_2$ and $u_1\,x$ is recursively convertible to $u'_2$.  We
then write $\WTEGCONV{t_1}{t_2}$.

The convertibility relation allows introducing a new typing rule
which says that two convertible well-formed types have the same
inhabitants.

\section[Subtyping rules]{Subtyping rules\index{Subtyping rules}
\label{subtyping-rules}}

At the moment, we did not take into account one rule between universes
which says that any term in a universe of index $i$ is also a term in
the universe of index $i+1$ (this is the {\em cumulativity} rule of
{\CIC}). This property extends the equivalence relation of
convertibility into a {\em subtyping} relation inductively defined by:
\begin{enumerate}
\item if $\WTEGCONV{t}{u}$ then $\WTEGLECONV{t}{u}$,
\item if $i \leq j$ then $\WTEGLECONV{\Type(i)}{\Type(j)}$,
\item for any $i$, $\WTEGLECONV{\Set}{\Type(i)}$,
\item $\WTEGLECONV{\Prop}{\Set}$, hence, by transitivity,
  $\WTEGLECONV{\Prop}{\Type(i)}$, for any $i$
\item if $\WTEGCONV{T}{U}$ and $\WTELECONV{\Gamma::(x:T)}{T'}{U'}$ then $\WTEGLECONV{\forall~x:T, T'}{\forall~x:U, U'}$.
\end{enumerate}

The conversion rule up to subtyping is now exactly:

\begin{description}\label{Conv}
\item[Conv]\index{Typing rules!Conv}
 \inference{
      \frac{\WTEG{U}{s}~~~~\WTEG{t}{T}~~~~\WTEGLECONV{T}{U}}{\WTEG{t}{U}}}
  \end{description}


\paragraph[Normal form.]{Normal form.\index{Normal form}\label{Normal-form}\label{Head-normal-form}\index{Head normal form}}
A term which cannot be any more reduced is said to be in {\em normal
  form}. There are several ways (or strategies) to apply the reduction
rules. Among them, we have to mention the {\em head reduction} which
will play an important role (see Chapter~\ref{Tactics}). Any term can
be written as $\lb x_1:T_1\mto \ldots \lb x_k:T_k \mto
(t_0\ t_1\ldots t_n)$ where
$t_0$ is not an application. We say then that $t_0$ is the {\em head
  of $t$}. If we assume that $t_0$ is $\lb x:T\mto u_0$ then one step of
$\beta$-head reduction of $t$ is:
\[\lb x_1:T_1\mto \ldots \lb x_k:T_k\mto (\lb x:T\mto u_0\ t_1\ldots t_n)
~\triangleright ~ \lb (x_1:T_1)\ldots(x_k:T_k)\mto
(\subst{u_0}{x}{t_1}\ t_2 \ldots t_n)\]
Iterating the process of head reduction until the head of the reduced
term is no more an abstraction leads to the {\em $\beta$-head normal
  form} of $t$:
\[ t \triangleright \ldots \triangleright
\lb x_1:T_1\mto \ldots\lb x_k:T_k\mto (v\ u_1
\ldots u_m)\]
where $v$ is not an abstraction (nor an application).  Note that the
head normal form must not be confused with the normal form since some
$u_i$ can be reducible.
%
Similar notions of head-normal forms involving $\delta$, $\iota$ and $\zeta$
reductions or any combination of those can also be defined.

\section[Inductive definitions]{Inductive Definitions\label{Cic-inductive-definitions}}

% Here we assume that the reader knows what is an inductive definition.

Formally, we can represent any {\em inductive definition\index{definition!inductive}} as \Ind{}{p}{\Gamma_I}{\Gamma_C} where:
\begin{itemize}
  \item $\Gamma_I$ determines the names and types of inductive types;
  \item $\Gamma_C$ determines the names and types of constructors of these inductive types;
  \item $p$ determines the number of parameters of these inductive types.
\end{itemize}
These inductive definitions, together with global assumptions and global definitions, then form the global environment.
%
Additionally, for any $p$ there always exists $\Gamma_P=[a_1:A_1;\dots;a_p:A_p]$
such that each $(t:T)\in\Gamma_I\cup\Gamma_C$ can be written as:
$\forall\Gamma_P, T^\prime$ where $\Gamma_P$ is called the {\em context of parameters\index{context of parameters}}.

\paragraph{Examples}

    \newcommand\ind[3]{$\mathsf{Ind}~[#1]\left(\hskip-.4em
    \begin{array}{r @{\mathrm{~:=~}} l}
      #2 & #3 \\
    \end{array}
    \hskip-.4em
    \right)$}
    \def\colon{@{\hskip.5em:\hskip.5em}}

The declaration for parameterized lists is:
\begin{latexonly}
    \vskip.5em

\ind{1}{[\List:\Set\ra\Set]}{\left[\begin{array}{r \colon l}
                                      \Nil & \forall A:\Set,\List~A \\
                                      \cons & \forall A:\Set, A \ra \List~A \ra \List~A
                                   \end{array}
                             \right]}
    \vskip.5em
\end{latexonly}
\begin{rawhtml}<pre><table style="border-spacing:0">
  <tr style="vertical-align:middle">
    <td style="width:10pt;text-align:center;font-family:sans-serif;font-style:italic">Ind</td>
    <td style="width:20pt;text-align:center">[1]</td>
    <td style="width:5pt;text-align:center">⎛<br>⎝</td>
    <td style="width:120pt;text-align:center">[ <span style="font-family:monospace">list : Set → Set</span> ]</td>
    <td style="width:20pt;text-align:center;font-family:monospace">:=</td>
    <td style="width:10pt;text-align:center">⎡<br>⎣</td>
    <td>
      <table style="border-spacing:0">
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">nil</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:=</td>
          <td style="text-align:left;font-family:monospace">∀A : Set, list A</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">cons</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:=</td>
          <td style="text-align:left;font-family:monospace">∀A : Set, A → list A → list A</td>
        </tr>
      </table>
    </td>
    <td style="width:10pt;text-align:center">⎤<br>⎦</td>
    <td style="width:5pt;text-align:center">⎞<br>⎠</td>
  </tr>
</table></pre>
\end{rawhtml}
\noindent which corresponds to the result of the \Coq\ declaration:
\begin{coq_example*}
Inductive list (A:Set) : Set :=
  | nil : list A
  | cons : A -> list A -> list A.
\end{coq_example*}

\noindent The declaration for a mutual inductive definition of {\tree} and {\forest} is:
\begin{latexonly}
    \vskip.5em
\ind{~}{\left[\begin{array}{r \colon l}\tree&\Set\\\forest&\Set\end{array}\right]}
       {\left[\begin{array}{r \colon l}
             \node & \forest \ra \tree\\
             \emptyf & \forest\\
             \consf & \tree \ra \forest \ra \forest\\
                       \end{array}\right]}
    \vskip.5em
\end{latexonly}
\begin{rawhtml}<pre><table style="border-spacing:0">
  <tr style="vertical-align:middle">
    <td style="width:10pt;text-align:center;font-family:sans-serif;font-style:italic">Ind</td>
    <td style="width:20pt;text-align:center">[1]</td>
    <td style="width:5pt;text-align:center">⎛<br>⎜<br>⎝</td>
    <td style="width:10pt;text-align:center">⎡<br>⎣</td>
    <td>
      <table style="border-spacing:0">
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">tree</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">Set</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">forest</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">Set</td>
        </tr>
      </table>
    </td>
    <td style="width:10pt;text-align:center">⎤<br>⎦</td>
    <td style="width:20pt;text-align:center;font-family:monospace">:=</td>
    <td style="width:10pt;text-align:center">⎡<br>⎢<br>⎣</td>
    <td>
      <table style="border-spacing:0">
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">node</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">forest → tree</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">emptyf</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">forest</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">consf</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">tree → forest → forest</td>
        </tr>
      </table>
    </td>
    <td style="width:10pt;text-align:center">⎤<br>⎥<br>⎦</td>
    <td style="width:5pt;text-align:center">⎞<br>⎟<br>⎠</td>
  </tr>
</table></pre>
\end{rawhtml}
\noindent which corresponds to the result of the \Coq\
declaration:
\begin{coq_example*}
Inductive tree : Set :=
    node : forest -> tree
with forest : Set :=
  | emptyf : forest
  | consf : tree -> forest -> forest.
\end{coq_example*}

\noindent The declaration for a mutual inductive definition of {\even} and {\odd} is:
\begin{latexonly}
    \newcommand\GammaI{\left[\begin{array}{r \colon l}
                        \even & \nat\ra\Prop \\
                        \odd & \nat\ra\Prop
                      \end{array}
                \right]}
    \newcommand\GammaC{\left[\begin{array}{r \colon l}
                        \evenO & \even~\nO \\
                        \evenS & \forall n : \nat, \odd~n \ra \even~(\nS~n)\\
                        \oddS & \forall n : \nat, \even~n \ra \odd~(\nS~n)
                      \end{array}
                \right]}
    \vskip.5em
    \ind{1}{\GammaI}{\GammaC}
    \vskip.5em
\end{latexonly}
\begin{rawhtml}<pre><table style="border-spacing:0">
  <tr style="vertical-align:middle">
    <td style="width:10pt;text-align:center;font-family:sans-serif;font-style:italic">Ind</td>
    <td style="width:20pt;text-align:center">[1]</td>
    <td style="width:5pt;text-align:center">⎛<br>⎜<br>⎝</td>
    <td style="width:10pt;text-align:center">⎡<br>⎣</td>
    <td>
      <table style="border-spacing:0">
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">even</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">nat → Prop</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">odd</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">nat → Prop</td>
        </tr>
      </table>
    </td>
    <td style="width:10pt;text-align:center">⎤<br>⎦</td>
    <td style="width:20pt;text-align:center;font-family:monospace">:=</td>
    <td style="width:10pt;text-align:center">⎡<br>⎢<br>⎣</td>
    <td>
      <table style="border-spacing:0">
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">even_O</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">even O</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">even_S</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">∀n : nat, odd n → even (S n)</td>
        </tr>
        <tr>
          <td style="width:20pt;text-align:right;font-family:monospace">odd_S</td>
          <td style="width:20pt;text-align:center;font-family:monospace">:</td>
          <td style="text-align:left;font-family:monospace">∀n : nat, even n → odd (S n)</td>
        </tr>
      </table>
    </td>
    <td style="width:10pt;text-align:center">⎤<br>⎥<br>⎦</td>
    <td style="width:5pt;text-align:center">⎞<br>⎟<br>⎠</td>
  </tr>
</table></pre>
\end{rawhtml}
\noindent which corresponds to the result of the \Coq\
declaration:
\begin{coq_example*}
Inductive even : nat -> Prop :=
  | even_O : even 0
  | even_S : forall n, odd n -> even (S n)
with odd : nat -> Prop :=
  | odd_S : forall n, even n -> odd (S n).
\end{coq_example*}

\subsection{Types of inductive objects}
We have to give the type of constants in a global environment $E$ which
contains an inductive declaration.

\begin{description}
\item[Ind]  \index{Typing rules!Ind}
  \inference{\frac{\WFE{\Gamma}~~~~~~~~\Ind{}{p}{\Gamma_I}{\Gamma_C} \in E~~~~~~~~(a:A)\in\Gamma_I}{\WTEG{a}{A}}}
\item[Constr]  \index{Typing rules!Constr}
  \inference{\frac{\WFE{\Gamma}~~~~~~~~\Ind{}{p}{\Gamma_I}{\Gamma_C} \in E~~~~~~~~(c:C)\in\Gamma_C}{\WTEG{c}{C}}}
\end{description}

\begin{latexonly}%
\paragraph{Example.}
Provided that our environment $E$ contains inductive definitions we showed before,
these two inference rules above enable us to conclude that:
\vskip.5em
\newcommand\prefix{E[\Gamma]\vdash\hskip.25em}
$\begin{array}{@{} l}
   \prefix\even : \nat\ra\Prop\\
   \prefix\odd : \nat\ra\Prop\\
   \prefix\evenO : \even~\nO\\
   \prefix\evenS : \forall~n:\nat, \odd~n \ra \even~(\nS~n)\\
   \prefix\oddS : \forall~n:\nat, \even~n \ra \odd~(\nS~n)
 \end{array}$
\end{latexonly}%

%\paragraph{Parameters.}
%%The parameters introduce a distortion between the inside specification
%%of the inductive declaration where parameters are supposed to be
%%instantiated (this representation is appropriate for checking the
%%correctness or deriving the destructor principle) and the outside
%%typing rules where the inductive objects are seen as objects
%%abstracted with respect to the parameters.

%In the definition of \List\ or \haslength\, $A$ is a parameter because
%what is effectively inductively defined is $\ListA$ or $\haslengthA$ for
%a given $A$ which is constant in the type of constructors.  But when
%we define $(\haslengthA~l~n)$, $l$ and $n$ are not parameters because the
%constructors manipulate different instances of this family.

\subsection{Well-formed inductive definitions}
We cannot accept any inductive declaration because some of them lead
to inconsistent systems.
We restrict ourselves to definitions which
satisfy a syntactic criterion of positivity. Before giving the formal
rules, we need a few definitions:

\paragraph[Definition]{Definition\index{Arity}\label{Arity}}
A type $T$ is an {\em arity of sort $s$} if it converts
to the sort $s$ or to a product $\forall~x:T,U$ with $U$ an arity
of sort $s$.

\paragraph[Examples]{Examples}
$A\ra \Set$ is an arity of sort $\Set$.
$\forall~A:\Prop,A\ra \Prop$ is an arity of sort \Prop.

\paragraph[Definition]{Definition}
A type $T$ is an {\em arity} if there is a $s\in\Sort$
such that $T$ is an arity of sort $s$.

\paragraph[Examples]{Examples}
$A\ra \Set$ and $\forall~A:\Prop,A\ra \Prop$ are arities.

\paragraph[Definition]{Definition\index{type of constructor}}
We say that $T$ is a {\em type of constructor of $I$\index{type of constructor}}
in one of the following two cases:
\begin{itemize}
  \item $T$ is $(I~t_1\ldots ~t_n)$
  \item $T$ is $\forall x:U,T^\prime$ where $T^\prime$ is also a type of constructor of $I$
\end{itemize}

\paragraph[Examples]{Examples}
$\nat$ and $\nat\ra\nat$ are types of constructors of $\nat$.\\
$\forall A:\Type,\List~A$ and $\forall A:\Type,A\ra\List~A\ra\List~A$ are constructors of $\List$.

\paragraph[Definition]{Definition\index{Positivity}\label{Positivity}}
The type of constructor $T$ will be said to {\em satisfy the positivity
condition} for a constant $X$ in the following cases:

\begin{itemize}
\item $T=(X~t_1\ldots ~t_n)$ and $X$ does not occur free in
any $t_i$
\item $T=\forall~x:U,V$ and $X$ occurs only strictly positively in $U$ and
the type $V$ satisfies the positivity condition for $X$
\end{itemize}
%
The constant $X$ {\em occurs strictly positively} in $T$ in the
following cases:
%
\begin{itemize}
\item $X$ does not occur in $T$
\item $T$ converts to $(X~t_1 \ldots ~t_n)$ and $X$ does not occur in
  any of $t_i$
\item $T$ converts to $\forall~x:U,V$ and $X$ does not occur in
  type $U$ but occurs strictly positively in type $V$
\item $T$ converts to $(I~a_1 \ldots ~a_m ~ t_1 \ldots ~t_p)$ where
  $I$ is the name of an inductive declaration of the form
  $\Ind{\Gamma}{m}{I:A}{c_1:\forall p_1:P_1,\ldots \forall
    p_m:P_m,C_1;\ldots;c_n:\forall p_1:P_1,\ldots \forall
    p_m:P_m,C_n}$ 
  (in particular, it is not mutually defined and it has $m$
  parameters) and $X$ does not occur in any of the $t_i$, and the
  (instantiated) types of constructor $C_i\{p_j/a_j\}_{j=1\ldots m}$
  of $I$ satisfy 
  the nested positivity condition for $X$
%\item more generally, when $T$ is not a type, $X$ occurs strictly
%positively in $T[x:U]u$ if $X$ does not occur in $U$ but occurs
%strictly positively in $u$
\end{itemize}
%
The type of constructor $T$ of $I$ {\em satisfies the nested
positivity condition} for a constant $X$ in the following
cases:

\begin{itemize}
\item $T=(I~b_1\ldots b_m~u_1\ldots ~u_{p})$, $I$ is an inductive
  definition with $m$ parameters and $X$ does not occur in
any $u_i$
\item $T=\forall~x:U,V$ and $X$ occurs only strictly positively in $U$ and
the type $V$ satisfies the nested positivity condition for $X$
\end{itemize}

\newcommand\vv{\textSFxi}    % │
\newcommand\hh{\textSFx}     %   ─
\newcommand\vh{\textSFviii}  % ├
\newcommand\hv{\textSFii}    %   └
\newlength\framecharacterwidth
\settowidth\framecharacterwidth{\hh}
\newcommand\ws{\hbox{}\hskip\the\framecharacterwidth}
\newcommand\ruleref[1]{\hskip.25em\dots\hskip.2em{\em (bullet #1)}}

\noindent For instance, if one considers the type

\begin{verbatim}
Inductive tree (A:Type) : Type :=
 | leaf : list A
 | node : A -> (nat -> tree A) -> tree A
\end{verbatim}

\begin{latexonly}
\noindent Then every instantiated constructor of $\ListA$ satisfies the nested positivity condition for $\List$\\
\noindent
\ws\ws\vv\\
\ws\ws\vh\hh\ws concerning type $\ListA$ of constructor $\Nil$:\\
\ws\ws\vv\ws\ws\ws\ws Type $\ListA$ of constructor $\Nil$ satisfies the positivity condition for $\List$\\
\ws\ws\vv\ws\ws\ws\ws because $\List$ does not appear in any (real) arguments of the type of that constructor\\
\ws\ws\vv\ws\ws\ws\ws (primarily because $\List$ does not have any (real) arguments)\ruleref1\\
\ws\ws\vv\\
\ws\ws\hv\hh\ws concerning type $\forall~A\ra\ListA\ra\ListA$ of constructor $\cons$:\\
\ws\ws\ws\ws\ws\ws\ws Type $\forall~A:\Type,A\ra\ListA\ra\ListA$ of constructor $\cons$\\
\ws\ws\ws\ws\ws\ws\ws satisfies the positivity condition for $\List$ because:\\
\ws\ws\ws\ws\ws\ws\ws\vv\\
\ws\ws\ws\ws\ws\ws\ws\vh\hh\ws $\List$ occurs only strictly positively in $\Type$\ruleref3\\
\ws\ws\ws\ws\ws\ws\ws\vv\\
\ws\ws\ws\ws\ws\ws\ws\vh\hh\ws $\List$ occurs only strictly positively in $A$\ruleref3\\
\ws\ws\ws\ws\ws\ws\ws\vv\\
\ws\ws\ws\ws\ws\ws\ws\vh\hh\ws $\List$ occurs only strictly positively in $\ListA$\ruleref4\\
\ws\ws\ws\ws\ws\ws\ws\vv\\
\ws\ws\ws\ws\ws\ws\ws\hv\hh\ws $\List$ satisfies the positivity condition for $\ListA$\ruleref1
\end{latexonly}
\begin{rawhtml}
<pre>
<span style="font-family:serif">Then every instantiated constructor of <span style="font-family:monospace">list A</span> satisfies the nested positivity condition for <span style="font-family:monospace">list</span></span>
  │
  ├─ <span style="font-family:serif">concerning type <span style="font-family:monospace">list A</span> of constructor <span style="font-family:monospace">nil</span>:</span>
  │    <span style="font-family:serif">Type <span style="font-family:monospace">list A</span> of constructor <span style="font-family:monospace">nil</span> satisfies the positivity condition for <span style="font-family:monospace">list</span></span>
  │    <span style="font-family:serif">because <span style="font-family:monospace">list</span> does not appear in any (real) arguments of the type of that constructor</span>
  │    <span style="font-family:serif">(primarily because list does not have any (real) arguments) ... <span style="font-style:italic">(bullet 1)</span></span>
  │
  ╰─ <span style="font-family:serif">concerning type <span style="font-family:monospace">∀ A → list A → list A</span> of constructor <span style="font-family:monospace">cons</span>:</span>
       <span style="font-family:serif">Type <span style="font-family:monospace">∀ A : Type, A → list A → list A</span> of constructor <span style="font-family:monospace">cons</span></span>
       <span style="font-family:serif">satisfies the positivity condition for <span style="font-family:monospace">list</span> because:</span>
       │
       ├─ <span style="font-family:serif"><span style="font-family:monospace">list</span> occurs only strictly positively in <span style="font-family:monospace">Type</span> ... <span style="font-style:italic">(bullet 3)</span></span>
       │
       ├─ <span style="font-family:serif"><span style="font-family:monospace">list</span> occurs only strictly positively in <span style="font-family:monospace">A</span> ... <span style="font-style:italic">(bullet 3)</span></span>
       │
       ├─ <span style="font-family:serif"><span style="font-family:monospace">list</span> occurs only strictly positively in <span style="font-family:monospace">list A</span> ... <span style="font-style:italic">(bullet 4)</span></span>
       │
       ╰─ <span style="font-family:serif"><span style="font-family:monospace">list</span> satisfies the positivity condition for <span style="font-family:monospace">list A</span> ... <span style="font-style:italic">(bullet 1)</span></span>
</pre>
\end{rawhtml}

\paragraph{Correctness rules.}
We shall now describe the rules allowing the introduction of a new
inductive definition.

\begin{description}
\item[W-Ind] Let $E$ be a global environment and
  $\Gamma_P,\Gamma_I,\Gamma_C$ are contexts such that
  $\Gamma_I$ is $[I_1:\forall \Gamma_P,A_1;\ldots;I_k:\forall
  \Gamma_P,A_k]$ and $\Gamma_C$ is 
  $[c_1:\forall \Gamma_P,C_1;\ldots;c_n:\forall \Gamma_P,C_n]$. 
\inference{
  \frac{
  (\WTE{\Gamma_P}{A_j}{s'_j})_{j=1\ldots  k}
  ~~~~~~~~ (\WTE{\Gamma_I;\Gamma_P}{C_i}{s_{q_i}})_{i=1\ldots  n}
}
  {\WF{E;\Ind{}{p}{\Gamma_I}{\Gamma_C}}{\Gamma}}}
provided that the following side conditions hold:
\begin{itemize}
\item $k>0$ and all of $I_j$ and $c_i$ are distinct names for $j=1\ldots  k$ and $i=1\ldots  n$,
\item $p$ is the number of parameters of \NInd{}{\Gamma_I}{\Gamma_C}
  and $\Gamma_P$ is the context of parameters, 
\item for $j=1\ldots  k$ we have that $A_j$ is an arity of sort $s_j$ and $I_j
  \notin E$,
\item for $i=1\ldots  n$ we have that $C_i$ is a type of constructor of
  $I_{q_i}$ which satisfies the positivity condition for $I_1 \ldots  I_k$
  and $c_i \notin \Gamma \cup E$.
\end{itemize}
\end{description}
One can remark that there is a constraint between the sort of the
arity of the inductive type and the sort of the type of its
constructors which will always be satisfied for the impredicative sort
{\Prop} but may fail to define inductive definition 
on sort \Set{} and generate constraints between universes for
inductive definitions in the {\Type} hierarchy.

\paragraph{Examples.}
It is well known that existential quantifier can be encoded as an
inductive definition.
The following declaration introduces the second-order existential
quantifier $\exists X.P(X)$.
\begin{coq_example*}
Inductive exProp (P:Prop->Prop) : Prop :=
  exP_intro : forall X:Prop, P X -> exProp P.
\end{coq_example*}
The same definition on \Set{} is not allowed and fails:
% (********** The following is not correct and should produce **********)
% (*** Error: Large non-propositional inductive types must be in Type***)
\begin{coq_example}
Fail Inductive exSet (P:Set->Prop) : Set :=
  exS_intro : forall X:Set, P X -> exSet P.
\end{coq_example}
It is possible to declare the same inductive definition in the
universe \Type. 
The \texttt{exType} inductive definition has type  $(\Type_i \ra\Prop)\ra
\Type_j$ with the constraint that the parameter \texttt{X} of \texttt{exT\_intro} has type $\Type_k$ with $k<j$ and $k\leq i$.
\begin{coq_example*}
Inductive exType (P:Type->Prop) : Type :=
  exT_intro : forall X:Type, P X -> exType P.
\end{coq_example*}
%We shall assume for the following definitions that, if necessary, we
%annotated the type of constructors such that we know if the argument
%is recursive or not.  We shall write the type $(x:_R T)C$ if it is 
%a recursive argument and $(x:_P T)C$ if the argument is not recursive.

\paragraph[Template polymorphism.]{Template polymorphism.\index{Template polymorphism}}
\label{Template-polymorphism}

Inductive types declared in {\Type} are
polymorphic over their arguments in {\Type}.
If $A$ is an arity of some sort and $s$ is a sort, we write $A_{/s}$ for the arity
obtained from $A$ by replacing its sort with $s$. Especially, if $A$
is well-typed in some global environment and local context, then $A_{/s}$ is typable
by typability of all products in the Calculus of Inductive Constructions.
The following typing rule is added to the theory.

\begin{description}
\item[Ind-Family] Let $\Ind{}{p}{\Gamma_I}{\Gamma_C}$ be an
  inductive definition. Let $\Gamma_P = [p_1:P_1;\ldots;p_{p}:P_{p}]$
  be its context of parameters, $\Gamma_I = [I_1:\forall
    \Gamma_P,A_1;\ldots;I_k:\forall \Gamma_P,A_k]$ its context of
  definitions and $\Gamma_C = [c_1:\forall
    \Gamma_P,C_1;\ldots;c_n:\forall \Gamma_P,C_n]$ its context of
  constructors, with $c_i$ a constructor of $I_{q_i}$.

  Let $m \leq p$ be the length of the longest prefix of parameters
  such that the $m$ first arguments of all occurrences of all $I_j$ in
  all $C_k$ (even the occurrences in the hypotheses of $C_k$) are
  exactly applied to $p_1~\ldots~p_m$ ($m$ is the number of {\em
    recursively uniform parameters} and the $p-m$ remaining parameters
  are the {\em recursively non-uniform parameters}). Let $q_1$,
  \ldots, $q_r$, with $0\leq r\leq m$, be a (possibly) partial
  instantiation of the recursively uniform parameters of
  $\Gamma_P$. We have:

\inference{\frac
{\left\{\begin{array}{l}
\Ind{}{p}{\Gamma_I}{\Gamma_C} \in E\\
(E[] \vdash q_l : P'_l)_{l=1\ldots r}\\
(\WTELECONV{}{P'_l}{\subst{P_l}{p_u}{q_u}_{u=1\ldots l-1}})_{l=1\ldots r}\\
1 \leq j \leq k
\end{array}
\right.}
{E[] \vdash I_j\,q_1\,\ldots\,q_r:\forall [p_{r+1}:P_{r+1};\ldots;p_{p}:P_{p}], (A_j)_{/s_j}}
}

provided that the following side conditions hold:

\begin{itemize}
\item $\Gamma_{P'}$ is the context obtained from $\Gamma_P$ by
replacing each $P_l$ that is an arity with $P'_l$ for $1\leq l \leq r$ (notice that
$P_l$ arity implies $P'_l$ arity since $\WTELECONV{}{P'_l}{ \subst{P_l}{p_u}{q_u}_{u=1\ldots l-1}}$);
\item there are sorts $s_i$, for $1 \leq i \leq k$ such that, for
 $\Gamma_{I'} = [I_1:\forall
    \Gamma_{P'},(A_1)_{/s_1};\ldots;I_k:\forall \Gamma_{P'},(A_k)_{/s_k}]$
we have $(\WTE{\Gamma_{I'};\Gamma_{P'}}{C_i}{s_{q_i}})_{i=1\ldots  n}$;
\item the sorts $s_i$ are such that all eliminations, to {\Prop}, {\Set} and
  $\Type(j)$, are allowed (see Section~\ref{allowedeleminationofsorts}).
\end{itemize}
\end{description}
%
Notice that if $I_j\,q_1\,\ldots\,q_r$ is typable using the rules {\bf
Ind-Const} and {\bf App}, then it is typable using the rule {\bf
Ind-Family}. Conversely, the extended theory is not stronger than the
theory without {\bf Ind-Family}. We get an equiconsistency result by
mapping each $\Ind{}{p}{\Gamma_I}{\Gamma_C}$ occurring into a
given derivation into as many different inductive types and constructors
as the number of different (partial) replacements of sorts, needed for
this derivation, in the parameters that are arities (this is possible
because $\Ind{}{p}{\Gamma_I}{\Gamma_C}$ well-formed implies
that $\Ind{}{p}{\Gamma_{I'}}{\Gamma_{C'}}$ is well-formed and
has the same allowed eliminations, where
$\Gamma_{I'}$ is defined as above and $\Gamma_{C'} = [c_1:\forall
\Gamma_{P'},C_1;\ldots;c_n:\forall \Gamma_{P'},C_n]$). That is,
the changes in the types of each partial instance
$q_1\,\ldots\,q_r$ can be characterized by the ordered sets of arity
sorts among the types of parameters, and to each signature is
associated a new inductive definition with fresh names. Conversion is
preserved as any (partial) instance $I_j\,q_1\,\ldots\,q_r$ or
$C_i\,q_1\,\ldots\,q_r$ is mapped to the names chosen in the specific
instance of $\Ind{}{p}{\Gamma_I}{\Gamma_C}$.

\newcommand{\Single}{\mbox{\textsf{Set}}}

In practice, the rule {\bf Ind-Family} is used by {\Coq} only when all the
inductive types of the inductive definition are declared with an arity whose 
sort is in the $\Type$
hierarchy. Then, the polymorphism is over the parameters whose
type is an arity of sort in the {\Type} hierarchy. 
The sort $s_j$ are
chosen canonically so that each $s_j$ is minimal with respect to the
hierarchy ${\Prop}\subset{\Set_p}\subset\Type$ where $\Set_p$ is
predicative {\Set}.
%and ${\Prop_u}$ is the sort of small singleton
%inductive types (i.e. of inductive types with one single constructor
%and that contains either proofs or inhabitants of singleton types
%only). 
More precisely, an empty or small singleton inductive definition
(i.e. an inductive definition of which all inductive types are
singleton -- see paragraph~\ref{singleton}) is set in
{\Prop}, a small non-singleton inductive type is set in {\Set} (even
in case {\Set} is impredicative -- see Section~\ref{impredicativity}),
and otherwise in the {\Type} hierarchy.

Note that the side-condition about allowed elimination sorts in the
rule~{\bf Ind-Family} is just to avoid to recompute the allowed
elimination sorts at each instance of a pattern-matching (see
section~\ref{elimdep}).  
As an example, let us consider the following definition:
\begin{coq_example*}
Inductive option (A:Type) : Type := 
| None : option A 
| Some : A -> option A.
\end{coq_example*}
%
As the definition is set in the {\Type} hierarchy, it is used
polymorphically over its parameters whose types are arities of a sort
in the {\Type} hierarchy. Here, the parameter $A$ has this property,
hence, if \texttt{option} is applied to a type in {\Set}, the result is
in {\Set}. Note that if \texttt{option} is applied to a type in {\Prop},
then, the result is not set in \texttt{Prop} but in \texttt{Set}
still. This is because \texttt{option} is not a singleton type (see
section~\ref{singleton}) and it would lose the elimination to {\Set} and
{\Type} if set in {\Prop}.

\begin{coq_example}
Check (fun A:Set => option A).
Check (fun A:Prop => option A).
\end{coq_example}
%
Here is another example.
%
\begin{coq_example*}
Inductive prod (A B:Type) : Type := pair : A -> B -> prod A B.
\end{coq_example*}
%
As \texttt{prod} is a singleton type, it will be in {\Prop} if applied
twice to propositions, in {\Set} if applied twice to at least one type
in {\Set} and none in {\Type}, and in {\Type} otherwise. In all cases,
the three kind of eliminations schemes are allowed.

\begin{coq_example}
Check (fun A:Set => prod A).
Check (fun A:Prop => prod A A).
Check (fun (A:Prop) (B:Set) => prod A B).
Check (fun (A:Type) (B:Prop) => prod A B).
\end{coq_example}

\Rem Template polymorphism used to be called ``sort-polymorphism of
inductive types'' before universe polymorphism (see
Chapter~\ref{Universes-full}) was introduced.

\subsection{Destructors}
The specification of inductive definitions with arities and
constructors is quite natural.  But we still have to say how to use an
object in an inductive type.

This problem is rather delicate. There are actually several different
ways to do that. Some of them are logically equivalent but not always
equivalent from the computational point of view or from the user point
of view.

From the computational point of view, we want to be able to define a
function whose domain is an inductively defined type by using a
combination of case analysis over the possible constructors of the
object and recursion.

Because we need to keep a consistent theory and also we prefer to keep
a strongly normalizing reduction, we cannot accept any sort of
recursion (even terminating). So the basic idea is to restrict
ourselves to primitive recursive functions and functionals.

For instance, assuming a parameter $A:\Set$ exists in the local context, we
want to build a function \length\ of type $\ListA\ra \nat$ which
computes the length of the list, so such that $(\length~(\Nil~A)) = \nO$
and $(\length~(\cons~A~a~l)) = (\nS~(\length~l))$.  We want these
equalities to be recognized implicitly and taken into account in the
conversion rule.

From the logical point of view, we have built a type family by giving
a set of constructors.  We want to capture the fact that we do not
have any other way to build an object in this type. So when trying to
prove a property about an object $m$ in an inductive definition it is
enough to enumerate all the cases where $m$ starts with a different
constructor.

In case the inductive definition is effectively a recursive one, we
want to capture the extra property that we have built the smallest
fixed point of this recursive equation.  This says that we are only
manipulating finite objects. This analysis provides induction
principles.  
For instance, in order to prove $\forall l:\ListA,(\haslengthA~l~(\length~l))$
it is enough to prove:
%
\begin{itemize}
  \item $(\haslengthA~(\Nil~A)~(\length~(\Nil~A)))$
  \item $\forall a:A, \forall l:\ListA, (\haslengthA~l~(\length~l)) \ra\\
        \ra (\haslengthA~(\cons~A~a~l)~(\length~(\cons~A~a~l)))$
\end{itemize}
%
which given the conversion equalities satisfied by \length\ is the
same as proving:
%
\begin{itemize}
  \item $(\haslengthA~(\Nil~A)~\nO)$
  \item $\forall a:A, \forall l:\ListA, (\haslengthA~l~(\length~l)) \ra\\
        \ra (\haslengthA~(\cons~A~a~l)~(\nS~(\length~l)))$
\end{itemize}
%
One conceptually simple way to do that, following the basic scheme
proposed by Martin-L\"of in his Intuitionistic Type Theory, is to
introduce for each inductive definition an elimination operator. At
the logical level it is a proof of the usual induction principle and
at the computational level it implements a generic operator for doing
primitive recursion over the structure.

But this operator is rather tedious to implement and use. We choose in
this version of {\Coq} to factorize the operator for primitive recursion
into two more primitive operations as was first suggested by Th. Coquand
in~\cite{Coq92}.  One is the definition by pattern-matching. The second one is a definition by guarded fixpoints. 

\subsubsection[The {\tt match\ldots with \ldots end} construction.]{The {\tt match\ldots with \ldots end} construction.\label{Caseexpr}
\index{match@{\tt match\ldots with\ldots end}}}

The basic idea of this operator is that we have an object
$m$ in an inductive type $I$ and we want to prove a property
which possibly depends on $m$. For this, it is enough to prove the
property for $m = (c_i~u_1\ldots  u_{p_i})$ for each constructor of $I$.
The \Coq{} term for this proof will be written:
\[\kw{match}~m~\kw{with}~ (c_1~x_{11}~...~x_{1p_1}) \Ra f_1 ~|~\ldots~|~
  (c_n~x_{n1}~...~x_{np_n}) \Ra f_n~ \kw{end}\]
In this expression, if
$m$ eventually happens to evaluate to $(c_i~u_1\ldots u_{p_i})$ then
the expression will behave as specified in its $i$-th branch and
it will reduce to $f_i$ where the $x_{i1}$\ldots $x_{ip_i}$ are replaced
by the $u_1\ldots u_{p_i}$ according to the $\iota$-reduction.

Actually, for type-checking a \kw{match\ldots with\ldots end}
expression we also need to know the predicate $P$ to be proved by case
analysis. In the general case where $I$ is an inductively defined
$n$-ary relation, $P$ is a predicate over $n+1$ arguments: the $n$ first ones
correspond to the arguments of $I$ (parameters excluded), and the last
one corresponds to object $m$. \Coq{} can sometimes infer this
predicate but sometimes not. The concrete syntax for describing this
predicate uses the \kw{as\ldots in\ldots return} construction. For
instance, let us assume that $I$ is an unary predicate with one
parameter and one argument. The predicate is made explicit using the syntax:
\[\kw{match}~m~\kw{as}~ x~ \kw{in}~ I~\verb!_!~a~ \kw{return}~ P
 ~\kw{with}~ (c_1~x_{11}~...~x_{1p_1}) \Ra f_1 ~|~\ldots~|~
  (c_n~x_{n1}~...~x_{np_n}) \Ra f_n \kw{end}\]
The \kw{as} part can be omitted if either the result type does not
depend on $m$ (non-dependent elimination) or $m$ is a variable (in
this case, $m$ can occur in $P$ where it is considered a bound variable).
The \kw{in} part can be
omitted if the result type does not depend on the arguments of
$I$. Note that the arguments of $I$ corresponding to parameters
\emph{must} be \verb!_!, because the result type is not generalized to
all possible values of the parameters.
The other arguments of $I$
(sometimes called indices in the literature)
% NOTE: e.g. http://www.qatar.cmu.edu/~sacchini/papers/types08.pdf
have to be variables
($a$ above) and these variables can occur in $P$.
The expression after \kw{in}
must be seen as an \emph{inductive type pattern}. Notice that
expansion of implicit arguments and notations apply to this pattern.
%
For the purpose of presenting the inference rules, we use a more
compact notation:
\[ \Case{(\lb a x \mto P)}{m}{ \lb x_{11}~...~x_{1p_1} \mto f_1 ~|~\ldots~|~
  \lb x_{n1}...x_{np_n} \mto f_n}\]

%% CP 06/06 Obsolete avec la nouvelle syntaxe et incompatible avec la
%% presentation theorique qui suit
% \paragraph{Non-dependent elimination.}
%
% When defining a function of codomain $C$ by case analysis over an
% object in an inductive type $I$, we build an object of type $I
% \ra C$. The minimality principle on an inductively defined logical
% predicate $I$ of type $A \ra \Prop$ is often used to prove a property
% $\forall x:A,(I~x)\ra (C~x)$.  These are particular cases of the dependent
% principle that we stated before with a predicate which does not depend
% explicitly on the object in the inductive definition.

% For instance, a function testing whether a list is empty 
% can be
% defined as:
% \[\kw{fun} l:\ListA \Ra \kw{match}~l~\kw{with}~ \Nil \Ra \true~
% |~(\cons~a~m) \Ra \false \kw{end}\]
% represented by
% \[\lb l:\ListA \mto\Case{\bool}{l}{\true~ |~ \lb a~m,~\false}\]
%\noindent {\bf Remark. } 

% In the system \Coq\ the expression above, can be
% written without mentioning
% the dummy abstraction:
% \Case{\bool}{l}{\Nil~ \mbox{\tt =>}~\true~ |~ (\cons~a~m)~
%  \mbox{\tt =>}~ \false}

\paragraph[Allowed elimination sorts.]{Allowed elimination sorts.\index{Elimination sorts}}
\label{allowedeleminationofsorts}

An important question for building the typing rule for \kw{match} is
what can be the type of $\lb a x \mto P$ with respect to the type of $m$. If
$m:I$ and
$I:A$ and
$\lb a x \mto P : B$
then by \compat{I:A}{B} we mean that one can use $\lb a x \mto P$ with $m$ in the above
match-construct.

\paragraph{Notations.}
The \compat{I:A}{B} is defined as the smallest relation satisfying the
following rules:
We write \compat{I}{B} for \compat{I:A}{B} where $A$ is the type of
$I$.

The case of inductive definitions in sorts \Set\ or \Type{} is simple.
There is no restriction on the sort of the predicate to be
eliminated. 
%
\begin{description}
\item[Prod] \inference{\frac{\compat{(I~x):A'}{B'}}
                      {\compat{I:\forall x:A, A'}{\forall x:A, B'}}}
\item[{\Set} \& \Type] \inference{\frac{
    s_1 \in \{\Set,\Type(j)\}~~~~~~~~s_2 \in \Sort}{\compat{I:s_1}{I\ra s_2}}}
\end{description}
%
The case of Inductive definitions of sort \Prop{} is a bit more
complicated, because of our interpretation of this sort. The only
harmless allowed elimination, is the one when predicate $P$ is also of
sort \Prop.
\begin{description}
\item[\Prop] \inference{\compat{I:\Prop}{I\ra\Prop}}
\end{description}
\Prop{} is the type of logical propositions, the proofs of properties
$P$ in \Prop{} could not be used for computation and are consequently
ignored by the extraction mechanism.
Assume $A$ and $B$ are two propositions, and the logical disjunction
$A\vee B$ is defined inductively by:
\begin{coq_example*}
Inductive or (A B:Prop) : Prop :=
  or_introl : A -> or A B | or_intror : B -> or A B.
\end{coq_example*}
The following definition which computes a boolean value by case over
the proof of \texttt{or A B} is not accepted:
% (***************************************************************)
% (*** This example should fail with ``Incorrect elimination'' ***)
\begin{coq_example}
Fail Definition choice (A B: Prop) (x:or A B) :=
  match x with or_introl _ _ a => true | or_intror _ _ b => false end.
\end{coq_example}
From the computational point of view, the structure of the proof of
\texttt{(or A B)} in this term is needed for computing the boolean
value.

In general, if $I$ has type \Prop\ then $P$ cannot have type $I\ra
\Set$, because it will mean to build an informative proof of type
$(P~m)$ doing a case analysis over a non-computational object that
will disappear in the extracted program.  But the other way is safe
with respect to our interpretation we can have $I$ a computational
object and $P$ a non-computational one, it just corresponds to proving
a logical property of a computational object.

% Also if $I$ is in one of the sorts \{\Prop, \Set\}, one cannot in
% general allow an elimination over a bigger sort such as \Type.  But
% this operation is safe whenever $I$ is a {\em small inductive} type,
% which means that all the types of constructors of
% $I$ are small with the following definition:\\
% $(I~t_1\ldots t_s)$ is a {\em small type of constructor} and
% $\forall~x:T,C$ is a small type of constructor if $C$ is and if $T$
% has type \Prop\ or \Set.  \index{Small inductive type}

% We call this particular elimination which gives the possibility to
% compute a type by induction on the structure of a term, a {\em strong
%   elimination}\index{Strong elimination}.

In the same spirit, elimination on $P$ of type $I\ra
\Type$ cannot be allowed because it trivially implies the elimination
on $P$ of type $I\ra \Set$ by cumulativity. It also implies that there
are two proofs of the same property which are provably different,
contradicting the proof-irrelevance property which is sometimes a
useful axiom:
\begin{coq_example}
Axiom proof_irrelevance : forall (P : Prop) (x y : P), x=y.
\end{coq_example}
\begin{coq_eval}
Reset proof_irrelevance.
\end{coq_eval}
The elimination of an inductive definition of type \Prop\ on a
predicate $P$ of type $I\ra \Type$ leads to a paradox when applied to 
impredicative inductive definition like the second-order existential
quantifier \texttt{exProp} defined above, because it give access to
the two projections on this type.

%\paragraph{Warning: strong elimination}
%\index{Elimination!Strong elimination}
%In previous versions of Coq, for a small inductive definition, only the
%non-informative strong elimination on \Type\ was allowed, because
%strong elimination on \Typeset\ was not compatible with the current
%extraction procedure. In this version, strong elimination on \Typeset\
%is accepted but a dummy element is extracted from it and may generate
%problems if extracted terms are explicitly used such as in the 
%{\tt Program} tactic or when extracting ML programs.

\paragraph[Empty and singleton elimination]{Empty and singleton elimination\label{singleton}
\index{Elimination!Singleton elimination}
\index{Elimination!Empty elimination}}

There are special inductive definitions in \Prop\ for which more
eliminations are allowed. 
\begin{description}
\item[\Prop-extended] 
\inference{
   \frac{I \mbox{~is an empty or singleton
       definition}~~~s \in \Sort}{\compat{I:\Prop}{I\ra s}}
}
\end{description}
%
% A {\em singleton definition} has always an informative content,
% even if it is a proposition.
%
A {\em singleton
definition} has only one constructor and all the arguments of this
constructor have type \Prop. In that case, there is a canonical
way to interpret the informative extraction on an object in that type,
such that the elimination on any sort $s$ is legal.  Typical examples are
the conjunction of non-informative propositions and the equality. 
If there is an hypothesis $h:a=b$ in the local context, it can be used for
rewriting not only in logical propositions but also in any type.
% In that case, the term \verb!eq_rec! which was defined as an axiom, is
% now a term of the calculus.
\begin{coq_example}
Print eq_rec.
Extraction eq_rec.
\end{coq_example}
An empty definition has no constructors, in that case also,
elimination on any sort is allowed.

\paragraph{Type of branches.}
Let $c$ be a term of type $C$, we assume $C$ is a type of constructor
for an inductive type $I$. Let $P$ be a term that represents the
property to be proved.
We assume $r$ is the number of parameters and $p$ is the number of arguments.

We define a new type \CI{c:C}{P} which represents the type of the
branch corresponding to the $c:C$ constructor.
\[
\begin{array}{ll}
\CI{c:(I~p_1\ldots p_r\ t_1 \ldots t_p)}{P} &\equiv (P~t_1\ldots ~t_p~c) \\[2mm]
\CI{c:\forall~x:T,C}{P} &\equiv \forall~x:T,\CI{(c~x):C}{P} 
\end{array}
\]
We write \CI{c}{P} for \CI{c:C}{P} with $C$ the type of $c$.

\paragraph{Example.}
The following term in concrete syntax:
\begin{verbatim}
match t as l return P' with
| nil _ => t1
| cons _ hd tl => t2
end
\end{verbatim}
can be represented in abstract syntax as $$\Case{P}{t}{f_1\,|\,f_2}$$
where
\begin{eqnarray*}
  P & = & \lambda~l~.~P^\prime\\
  f_1 & = & t_1\\
  f_2 & = & \lambda~(hd:\nat)~.~\lambda~(tl:\List~\nat)~.~t_2
\end{eqnarray*}
According to the definition:
\begin{latexonly}\vskip.5em\noindent\end{latexonly}%
\begin{htmlonly}

\end{htmlonly}
$ \CI{(\Nil~\nat)}{P} \equiv \CI{(\Nil~\nat) : (\List~\nat)}{P} \equiv (P~(\Nil~\nat))$
\begin{latexonly}\vskip.5em\noindent\end{latexonly}%
\begin{htmlonly}

\end{htmlonly}
$ \CI{(\cons~\nat)}{P}
  \equiv\CI{(\cons~\nat) : (\nat\ra\List~\nat\ra\List~\nat)}{P} \equiv\\
  \equiv\forall n:\nat, \CI{(\cons~\nat~n) : \List~\nat\ra\List~\nat)}{P} \equiv\\
  \equiv\forall n:\nat, \forall l:\List~\nat, \CI{(\cons~\nat~n~l) : \List~\nat)}{P} \equiv\\
\equiv\forall n:\nat, \forall l:\List~\nat,(P~(\cons~\nat~n~l))$.
\begin{latexonly}\vskip.5em\noindent\end{latexonly}%
\begin{htmlonly}

\end{htmlonly}
Given some $P$, then \CI{(\Nil~\nat)}{P} represents the expected type of $f_1$, and
\CI{(\cons~\nat)}{P} represents the expected type of $f_2$.

\paragraph{Typing rule.}

Our very general destructor for inductive definition enjoys the
following typing rule
% , where we write 
% \[
% \Case{P}{c}{[x_{11}:T_{11}]\ldots[x_{1p_1}:T_{1p_1}]g_1\ldots
%   [x_{n1}:T_{n1}]\ldots[x_{np_n}:T_{np_n}]g_n}
% \]
% for 
% \[
% \Case{P}{c}{(c_1~x_{11}~...~x_{1p_1}) \Ra g_1 ~|~\ldots~|~
% (c_n~x_{n1}...x_{np_n}) \Ra g_n }
% \]

\begin{description}
\item[match] \label{elimdep} \index{Typing rules!match}
\inference{
\frac{\WTEG{c}{(I~q_1\ldots  q_r~t_1\ldots  t_s)}~~
  \WTEG{P}{B}~~\compat{(I~q_1\ldots  q_r)}{B}
 ~~
(\WTEG{f_i}{\CI{(c_{p_i}~q_1\ldots  q_r)}{P}})_{i=1\ldots  l}}
{\WTEG{\Case{P}{c}{f_1|\ldots  |f_l}}{(P\ t_1\ldots  t_s\ c)}}}%\\[3mm]

provided $I$ is an inductive type in a definition
\Ind{}{r}{\Gamma_I}{\Gamma_C} with 
$\Gamma_C = [c_1:C_1;\ldots;c_n:C_n]$ and $c_{p_1}\ldots  c_{p_l}$ are the
only constructors of $I$.
\end{description}

\paragraph{Example.}

Below is a typing rule for the term shown in the previous example:
\inference{
  \frac{%
    \WTEG{t}{(\List~\nat)}~~~~%
    \WTEG{P}{B}~~~~%
    \compat{(\List~\nat)}{B}~~~~%
    \WTEG{f_1}{\CI{(\Nil~\nat)}{P}}~~~~%
    \WTEG{f_2}{\CI{(\cons~\nat)}{P}}%
  }
{\WTEG{\Case{P}{t}{f_1|f_2}}{(P~t)}}}

\paragraph[Definition of $\iota$-reduction.]{Definition of $\iota$-reduction.\label{iotared}
\index{iota-reduction@$\iota$-reduction}}
We still have to define the $\iota$-reduction in the general case.

A $\iota$-redex is a term of the following form:
\[\Case{P}{(c_{p_i}~q_1\ldots  q_r~a_1\ldots  a_m)}{f_1|\ldots |
    f_l}\]
with $c_{p_i}$ the $i$-th constructor of the inductive type $I$ with $r$
parameters.

The $\iota$-contraction of this term is $(f_i~a_1\ldots a_m)$ leading
to the general reduction rule:
\[ \Case{P}{(c_{p_i}~q_1\ldots  q_r~a_1\ldots  a_m)}{f_1|\ldots |
    f_n} \triangleright_{\iota} (f_i~a_1\ldots a_m) \]

\subsection[Fixpoint definitions]{Fixpoint definitions\label{Fix-term} \index{Fix@{\tt Fix}}}
The second operator for elimination is fixpoint definition. 
This fixpoint may involve several mutually recursive definitions.
The basic concrete syntax for a recursive set of mutually recursive 
declarations is (with $\Gamma_i$ contexts): 
\[\kw{fix}~f_1 (\Gamma_1) :A_1:=t_1~\kw{with} \ldots \kw{with}~ f_n
(\Gamma_n) :A_n:=t_n\]
The terms are obtained by projections from this set of declarations
and are written 
\[\kw{fix}~f_1 (\Gamma_1) :A_1:=t_1~\kw{with} \ldots \kw{with}~ f_n
(\Gamma_n) :A_n:=t_n~\kw{for}~f_i\]
In the inference rules, we represent such a
term by 
\[\Fix{f_i}{f_1:A_1':=t_1' \ldots f_n:A_n':=t_n'}\]
with $t_i'$ (resp. $A_i'$) representing the term $t_i$ abstracted
(resp. generalized) with
respect to the bindings in the context $\Gamma_i$, namely
$t_i'=\lb \Gamma_i \mto t_i$ and $A_i'=\forall \Gamma_i, A_i$.

\subsubsection{Typing rule}
The typing rule is the expected one for a fixpoint.

\begin{description}
\item[Fix] \index{Typing rules!Fix}
\inference{\frac{(\WTEG{A_i}{s_i})_{i=1\ldots n}~~~~
            (\WTE{\Gamma,f_1:A_1,\ldots,f_n:A_n}{t_i}{A_i})_{i=1\ldots n}}
         {\WTEG{\Fix{f_i}{f_1:A_1:=t_1 \ldots f_n:A_n:=t_n}}{A_i}}}
\end{description}
%
Any fixpoint definition cannot be accepted because non-normalizing terms
allow proofs of absurdity.
%
The basic scheme of recursion that should be allowed is the one needed for 
defining primitive
recursive functionals. In that case the fixpoint enjoys a special
syntactic restriction, namely one of the arguments belongs to an
inductive type, the function starts with a case analysis and recursive
calls are done on variables coming from patterns and representing subterms.
%
For instance in the case of natural numbers, a proof of the induction
principle of type 
\[\forall P:\nat\ra\Prop, (P~\nO)\ra(\forall n:\nat, (P~n)\ra(P~(\nS~n)))\ra
\forall n:\nat, (P~n)\]
can be represented by the term:
\[\begin{array}{l}
\lb P:\nat\ra\Prop\mto\lb f:(P~\nO)\mto \lb g:(\forall n:\nat,
(P~n)\ra(P~(\nS~n))) \mto\\
\Fix{h}{h:\forall n:\nat, (P~n):=\lb n:\nat\mto \Case{P}{n}{f~|~\lb
    p:\nat\mto (g~p~(h~p))}}
\end{array}
\]
%
Before accepting a fixpoint definition as being correctly typed, we
check that the definition is ``guarded''. A precise analysis of this
notion can be found in~\cite{Gim94}.
%
The first stage is to precise on which argument the fixpoint will be
decreasing. The type of this argument should be an inductive
definition.
%
For doing this, the syntax of fixpoints is extended and becomes 
 \[\Fix{f_i}{f_1/k_1:A_1:=t_1 \ldots f_n/k_n:A_n:=t_n}\]
where $k_i$ are positive integers.
Each $k_i$ represents the index of pararameter of $f_i$, on which $f_i$ is decreasing.
Each $A_i$ should be a type (reducible to a term) starting with at least
$k_i$ products $\forall y_1:B_1,\ldots \forall y_{k_i}:B_{k_i}, A'_i$ 
and $B_{k_i}$ an is unductive type.

Now in the definition $t_i$, if $f_j$ occurs then it should be applied
to at least $k_j$ arguments and the $k_j$-th argument should be
syntactically recognized as structurally smaller than $y_{k_i}$


The definition of being structurally smaller is a bit technical.
One needs first to  define the notion of 
{\em recursive arguments of a constructor}\index{Recursive arguments}.
For an inductive definition \Ind{}{r}{\Gamma_I}{\Gamma_C},
if the type of a constructor $c$ has the form
$\forall p_1:P_1,\ldots \forall p_r:P_r, 
\forall x_1:T_1, \ldots \forall x_r:T_r, (I_j~p_1\ldots 
p_r~t_1\ldots  t_s)$, then the recursive arguments will correspond to $T_i$ in
which one of the $I_l$ occurs.

The main rules for being structurally smaller are the following:\\
Given a variable $y$ of type an inductive
definition in a declaration 
\Ind{}{r}{\Gamma_I}{\Gamma_C}
where $\Gamma_I$ is $[I_1:A_1;\ldots;I_k:A_k]$, and $\Gamma_C$ is
  $[c_1:C_1;\ldots;c_n:C_n]$.
The terms structurally smaller than $y$ are:
\begin{itemize}
\item $(t~u)$ and $\lb x:u \mto t$ when $t$ is structurally smaller than $y$.
\item \Case{P}{c}{f_1\ldots f_n} when each $f_i$ is structurally
  smaller than $y$. \\
  If $c$ is $y$ or is structurally smaller than $y$, its type is an inductive
  definition $I_p$ part of the inductive
  declaration corresponding to $y$. 
  Each $f_i$ corresponds to a type of constructor $C_q \equiv
  \forall p_1:P_1,\ldots,\forall p_r:P_r, \forall y_1:B_1, \ldots \forall y_k:B_k, (I~a_1\ldots a_k)$ 
  and can consequently be
  written $\lb y_1:B'_1\mto \ldots \lb y_k:B'_k\mto g_i$.
  ($B'_i$ is obtained from $B_i$ by substituting parameters variables)
  the variables $y_j$ occurring
  in $g_i$ corresponding to recursive arguments $B_i$ (the ones in
  which one of the $I_l$ occurs) are structurally smaller than $y$.
\end{itemize}
The following definitions are correct, we enter them using the
{\tt Fixpoint} command as described in Section~\ref{Fixpoint} and show
the internal representation.
\begin{coq_example}
Fixpoint plus (n m:nat) {struct n} : nat :=
  match n with
  | O => m
  | S p => S (plus p m)
  end.
Print plus.
Fixpoint lgth (A:Set) (l:list A) {struct l} : nat :=
  match l with
  | nil _ => O
  | cons _ a l' => S (lgth A l')
  end.
Print lgth.
Fixpoint sizet (t:tree) : nat := let (f) := t in S (sizef f)
 with sizef (f:forest) : nat :=
  match f with
  | emptyf => O
  | consf t f => plus (sizet t) (sizef f)
  end.
Print sizet.
\end{coq_example}


\subsubsection[Reduction rule]{Reduction rule\index{iota-reduction@$\iota$-reduction}}
Let $F$ be the set of declarations: $f_1/k_1:A_1:=t_1 \ldots
f_n/k_n:A_n:=t_n$.
The reduction for fixpoints is:
\[ (\Fix{f_i}{F}~a_1\ldots
a_{k_i}) \triangleright_{\iota} \substs{t_i}{f_k}{\Fix{f_k}{F}}{k=1\ldots n}
~a_1\ldots a_{k_i}\]
when $a_{k_i}$ starts with a constructor.
This last restriction is needed in order to keep strong normalization
and corresponds to the reduction for primitive recursive operators.
%
The following reductions are now possible:
\def\plus{\mathsf{plus}}
\def\tri{\triangleright_\iota}
\begin{eqnarray*}
  \plus~(\nS~(\nS~\nO))~(\nS~\nO) & \tri & \nS~(\plus~(\nS~\nO)~(\nS~\nO))\\
                                  & \tri & \nS~(\nS~(\plus~\nO~(\nS~\nO)))\\
                                  & \tri & \nS~(\nS~(\nS~\nO))\\
\end{eqnarray*}

% La disparition de Program devrait rendre la construction Match obsolete
% \subsubsection{The {\tt Match \ldots with \ldots end} expression}
% \label{Matchexpr}
% %\paragraph{A unary {\tt Match\ldots with \ldots end}.}
% \index{Match...with...end@{\tt Match \ldots with \ldots end}}
% The {\tt Match} operator which was a primitive notion in older
% presentations of the Calculus of Inductive Constructions is now just a
% macro definition which generates the good combination of {\tt Case}
% and {\tt Fix} operators in order to generate an operator for primitive
% recursive definitions. It always considers an inductive definition as 
% a single inductive definition.

% The following examples illustrates this feature.
% \begin{coq_example}
% Definition nat_pr : (C:Set)C->(nat->C->C)->nat->C 
%   :=[C,x,g,n]Match n with x g end.
% Print nat_pr.
% \end{coq_example}
% \begin{coq_example}
% Definition forest_pr 
%   : (C:Set)C->(tree->forest->C->C)->forest->C
%   := [C,x,g,n]Match n with x g end.
% \end{coq_example}

% Cet exemple faisait error (HH le 12/12/96), j'ai change pour une
% version plus simple
%\begin{coq_example}
%Definition forest_pr 
%  : (P:forest->Set)(P emptyf)->((t:tree)(f:forest)(P f)->(P (consf t f)))
%    ->(f:forest)(P f)
%  := [C,x,g,n]Match n with x g end.
%\end{coq_example}

\subsubsection{Mutual induction}

The principles of mutual induction can be automatically generated 
using the {\tt Scheme} command described in Section~\ref{Scheme}.

\section{Admissible rules for global environments}

From the original rules of the type system, one can show the
admissibility of rules which change the local context of definition of
objects in the global environment. We show here the admissible rules
that are used used in the discharge mechanism at the end of a section.

% This is obsolete: Abstraction over defined constants actually uses a
% let-in since there are let-ins in Coq

%% \paragraph{Mechanism of substitution.}

%% One rule which can be proved valid, is to replace a term $c$ by its
%% value in the global environment. As we defined the substitution of a term for
%% a variable in a term, one can define the substitution of a term for a
%% constant. One easily extends this substitution to local contexts and global
%% environments.

%% \paragraph{Substitution Property:}
%% \inference{\frac{\WF{E;c:=t:T; E'}{\Gamma}}
%%            {\WF{E; \subst{E'}{c}{t}}{\subst{\Gamma}{c}{t}}}}

\paragraph{Abstraction.}

One can modify a global declaration by generalizing it over a
previously assumed constant $c$. For doing that, we need to modify the
reference to the global declaration in the subsequent global
environment and local context by explicitly applying this constant to
the constant $c'$.

Below, if $\Gamma$ is a context of the form
$[y_1:A_1;\ldots;y_n:A_n]$, we write $\forall
x:U,\subst{\Gamma}{c}{x}$ to mean
$[y_1:\forall~x:U,\subst{A_1}{c}{x};\ldots;y_n:\forall~x:U,\subst{A_n}{c}{x}]$
and
$\subst{E}{|\Gamma|}{|\Gamma|c}$.
to mean the parallel substitution
$\subst{\subst{E}{y_1}{(y_1~c)}\ldots}{y_n}{(y_n~c)}$.

\paragraph{First abstracting property:}
 \inference{\frac{\WF{E;c:U;E';c':=t:T;E''}{\Gamma}}
                 {\WF{E;c:U;E';c':=\lb x:U\mto \subst{t}{c}{x}:\forall~x:U,\subst{T}{c}{x};
                      \subst{E''}{c'}{(c'~c)}}{\subst{\Gamma}{c}{(c~c')}}}}

 \inference{\frac{\WF{E;c:U;E';c':T;E''}{\Gamma}}
                 {\WF{E;c:U;E';c':\forall~x:U,\subst{T}{c}{x};
                      \subst{E''}{c'}{(c'~c)}}{\subst{\Gamma}{c}{(c~c')}}}}

 \inference{\frac{\WF{E;c:U;E';\Ind{}{p}{\Gamma_I}{\Gamma_C};E''}{\Gamma}}
                 {\WFTWOLINES{E;c:U;E';\Ind{}{p+1}{\forall x:U,\subst{\Gamma_I}{c}{x}}{\forall x:U,\subst{\Gamma_C}{c}{x}};\subst{E''}{|\Gamma_I,\Gamma_C|}{|\Gamma_I,\Gamma_C|~c}}{\subst{\Gamma}{|\Gamma_I,\Gamma_C|}{|\Gamma_I,\Gamma_C|~c}}}}
%
One can similarly modify a global declaration by generalizing it over
a previously defined constant~$c'$.  Below, if $\Gamma$ is a context
of the form $[y_1:A_1;\ldots;y_n:A_n]$, we write $
\subst{\Gamma}{c}{u}$ to mean
$[y_1:\subst{A_1}{c}{u};\ldots;y_n:\subst{A_n}{c}{u}]$.

\paragraph{Second abstracting property:}
 \inference{\frac{\WF{E;c:=u:U;E';c':=t:T;E''}{\Gamma}}
           {\WF{E;c:=u:U;E';c':=(\letin{x}{u:U}{\subst{t}{c}{x}}):\subst{T}{c}{u};E''}{\Gamma}}}

 \inference{\frac{\WF{E;c:=u:U;E';c':T;E''}{\Gamma}}
           {\WF{E;c:=u:U;E';c':\subst{T}{c}{u};E''}{\Gamma}}}

 \inference{\frac{\WF{E;c:=u:U;E';\Ind{}{p}{\Gamma_I}{\Gamma_C};E''}{\Gamma}}
                 {\WF{E;c:=u:U;E';\Ind{}{p}{\subst{\Gamma_I}{c}{u}}{\subst{\Gamma_C}{c}{u}};E''}{\Gamma}}}

\paragraph{Pruning the local context.}
If one abstracts or substitutes constants with the above rules then it
may happen that some declared or defined constant does not occur any
more in the subsequent global environment and in the local context. One can
consequently derive the following property.

\paragraph{First pruning property:}
\inference{\frac{\WF{E;c:U;E'}{\Gamma} \qquad c \mbox{ does not occur in $E'$ and $\Gamma$}}
                {\WF{E;E'}{\Gamma}}}
 
\paragraph{Second pruning property:}
\inference{\frac{\WF{E;c:=u:U;E'}{\Gamma} \qquad c \mbox{ does not occur in $E'$ and $\Gamma$}}
                {\WF{E;E'}{\Gamma}}}
 
\section{Co-inductive types}
The implementation contains also co-inductive definitions, which are
types inhabited by infinite objects. 
More information on co-inductive definitions can be found
in~\cite{Gimenez95b,Gim98,GimCas05}.
%They are described in Chapter~\ref{Co-inductives}.

\section[The Calculus of Inductive Construction with
  impredicative \Set]{The Calculus of Inductive Construction with
  impredicative \Set\label{impredicativity}}

\Coq{} can be used as a type-checker for the
Calculus of Inductive Constructions with an impredicative sort \Set{}
by using the compiler option \texttt{-impredicative-set}.
%
For example, using the ordinary \texttt{coqtop} command, the following
is rejected.
% (** This example should fail *******************************
%     Error: The term forall X:Set, X -> X has type Type
%     while it is expected to have type Set ***)
\begin{coq_example}
Fail Definition id: Set := forall X:Set,X->X.
\end{coq_example}
while it will type-check, if one uses instead the \texttt{coqtop
  -impredicative-set} command.

The major change in the theory concerns the rule for product formation
in the sort \Set, which is extended to a domain in any sort:
\begin{description}
\item [Prod]  \index{Typing rules!Prod (impredicative Set)}
\inference{\frac{\WTEG{T}{s}~~~~s \in \Sort~~~~~~
    \WTE{\Gamma::(x:T)}{U}{\Set}}
      { \WTEG{\forall~x:T,U}{\Set}}} 
\end{description}
This extension has consequences on the inductive definitions which are
allowed. 
In the impredicative system, one can build so-called {\em large inductive
  definitions} like the example of second-order existential
quantifier (\texttt{exSet}).

There should be restrictions on the eliminations which can be
performed on such definitions. The eliminations rules in the
impredicative system for sort \Set{} become:
\begin{description}
\item[\Set] \inference{\frac{s \in
      \{\Prop, \Set\}}{\compat{I:\Set}{I\ra s}}
~~~~\frac{I \mbox{~is a small inductive definition}~~~~s \in
      \{\Type(i)\}}
         {\compat{I:\Set}{I\ra s}}}
\end{description}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Reference-Manual"
%%% End: 


