\chapter[Vernacular commands]{Vernacular commands\label{Vernacular-commands}
\label{Other-commands}}

\section{Displaying}

\subsection[\tt Print {\qualid}.]{\tt Print {\qualid}.\comindex{Print}}
This command displays on the screen informations about the declared or
defined object referred by {\qualid}.

\begin{ErrMsgs}
\item {\qualid} \errindex{not a defined object}
\end{ErrMsgs}

\begin{Variants}
\item {\tt Print Term {\qualid}.}
\comindex{Print Term}\\ 
This is a synonym to {\tt Print {\qualid}} when {\qualid} denotes a
global constant. 

\item {\tt About {\qualid}.}
\label{About}
\comindex{About}\\ 
This displays various informations about the object denoted by {\qualid}:
its kind (module, constant, assumption, inductive,
constructor, abbreviation\ldots), long name, type, implicit
arguments and argument scopes.

%\item {\tt Print Proof {\qualid}.}\comindex{Print Proof}\\
%In case \qualid\ denotes an opaque theorem defined in a section,
%it is stored on a special unprintable form and displayed as 
%{\tt <recipe>}. {\tt Print Proof} forces the printable form of \qualid\
%to be computed and displays it.
\end{Variants}

\subsection[\tt Print All.]{\tt Print All.\comindex{Print All}}
This command displays informations about the current state of the
environment, including sections and modules.

\begin{Variants}
\item {\tt Inspect \num.}\comindex{Inspect}\\
This command displays the {\num} last objects of the current
environment, including sections and modules.
\item {\tt Print Section {\ident}.}\comindex{Print Section}\\
should correspond to a currently open section, this command
displays the objects defined since the beginning of this section.
% Discontinued
%% \item {\tt Print.}\comindex{Print}\\
%% This command displays the axioms and variables declarations in the
%% environment as well as the constants defined since the last variable
%% was introduced.
\end{Variants}

\section{Requests to the environment}

\subsection[\tt Check {\term}.]{\tt Check {\term}.\label{Check}
\comindex{Check}}
This command displays the type of {\term}. When called in proof mode, 
the term is checked in the local context of the current subgoal.

\subsection[\tt Eval {\rm\sl convtactic} in {\term}.]{\tt Eval {\rm\sl convtactic} in {\term}.\comindex{Eval}}

This command performs the specified reduction on {\term}, and displays
the resulting term with its type. The term to be reduced may depend on
hypothesis introduced in the first subgoal (if a proof is in
progress).

\SeeAlso Section~\ref{Conversion-tactics}.

\subsection[\tt Extraction \term.]{\tt Extraction \term.\label{ExtractionTerm}
\comindex{Extraction}} 
This command displays the extracted term from
{\term}. The extraction is processed according to the distinction
between {\Set} and {\Prop}; that is to say, between logical and
computational content (see Section~\ref{Sorts}). The extracted term is
displayed in Objective Caml syntax, where global identifiers are still
displayed as in \Coq\ terms.

\begin{Variants}
\item \texttt{Recursive Extraction {\qualid$_1$} \ldots{} {\qualid$_n$}.}\\
  Recursively extracts all the material needed for the extraction of 
  globals {\qualid$_1$} \ldots{} {\qualid$_n$}.
\end{Variants}

\SeeAlso Chapter~\ref{Extraction}.

\subsection[\tt Opaque \qualid$_1$ \dots \qualid$_n$.]{\tt Opaque \qualid$_1$ \dots \qualid$_n$.\comindex{Opaque}\label{Opaque}} This command tells not to unfold the
the constants {\qualid$_1$} \dots {\qualid$_n$} in tactics using
$\delta$-conversion. Unfolding a constant is replacing it by its
definition. {\tt Opaque} can only apply on constants originally
defined as {\tt Transparent}.

Constants defined by a proof ended by {\tt Qed} are automatically
stamped as {\tt Opaque} and can no longer be considered as {\tt
Transparent}. This is to keep with the usual mathematical practice of
{\em proof irrelevance}: what matters in a mathematical development is
the sequence of lemma statements, not their actual proofs. This
distinguishes lemmas from the usual defined constants, whose actual
values are of course relevant in general.

\SeeAlso sections \ref{Conversion-tactics}, \ref{Automatizing},
\ref{Theorem}

\begin{ErrMsgs}
\item \errindex{The reference \qualid\ was not found in the current
environment}\\
    There is no constant referred by {\qualid} in the environment.
    Nevertheless, if you asked \texttt{Opaque foo bar}
    and if \texttt{bar} does not exist, \texttt{foo} is set opaque.
\end{ErrMsgs}

\subsection[\tt Transparent \qualid$_1$ \dots \qualid$_n$.]{\tt Transparent \qualid$_1$ \dots \qualid$_n$.\comindex{Transparent}\label{Transparent}}
This command is the converse of {\tt Opaque} and can only apply on constants originally defined as {\tt Transparent} to restore their initial behaviour after an {\tt Opaque} command.

The constants automatically declared transparent are the ones defined by a proof ended by {\tt Defined}, or by a {\tt
  Definition} or {\tt Local} with an explicit body.

\Warning {\tt Transparent} and \texttt{Opaque} are not synchronous
with the reset mechanism. If a constant was transparent at point A, if
you set it opaque at point B and reset to point A, you return to state
of point A with the difference that the constant is still opaque. This
can cause changes in tactic scripts behaviour.

At section or module closing, a constant recovers the status it got at
the time of its definition.

%TODO: expliquer le rapport avec les sections

\begin{ErrMsgs}
% \item \errindex{Can not set transparent.}\\
%     It is a constant from a required module or a parameter.
\item \errindex{The reference \qualid\ was not found in the current
environment}\\
    There is no constant referred by {\qualid} in the environment.
\end{ErrMsgs}

\SeeAlso sections \ref{Conversion-tactics}, \ref{Automatizing},
\ref{Theorem}

\subsection[\tt Search {\qualid}.]{\tt Search {\qualid}.\comindex{Search}}
This command displays the name and type of all theorems of the current
context whose statement's conclusion has the form {\tt ({\qualid} t1 ..
  tn)}.  This command is useful to remind the user of the name of
library lemmas.
\begin{ErrMsgs}
\item \errindex{The reference \qualid\ was not found in the current
environment}\\
    There is no constant in the environment named \qualid.
\end{ErrMsgs}

\begin{Variants}
\item
{\tt Search {\qualid} inside {\module$_1$} \ldots{} {\module$_n$}.}

This restricts the search to constructions defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\item {\tt Search {\qualid} outside {\module$_1$} \ldots{} {\module$_n$}.}

This restricts the search to constructions not defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\begin{ErrMsgs}
\item \errindex{Module/section \module{} not found}
No module \module{} has been required (see Section~\ref{Require}).
\end{ErrMsgs}

\end{Variants}

\subsection[\tt SearchAbout {\qualid}.]{\tt SearchAbout {\qualid}.\comindex{SearchAbout}}
This command displays the name and type of all objects (theorems,
axioms, etc) of the current context whose statement contains \qualid.
This command is useful to remind the user of the name of library
lemmas.

\begin{ErrMsgs}
\item \errindex{The reference \qualid\ was not found in the current
environment}\\
    There is no constant in the environment named \qualid.
\end{ErrMsgs}

\begin{Variants}
\item {\tt SearchAbout [ \nelist{\textrm{\textsl{qualid-or-string}}}{}
].}\\
\noindent where {\textrm{\textsl{qualid-or-string}}} is a {\qualid} or
a {\str}.

This extension of {\tt SearchAbout} searches for all objects whose
statement mentions all of {\qualid} of the list and whose name
contains all {\str} of the list.

\Example

\begin{coq_example}
Require Import ZArith.
SearchAbout [ Zmult Zplus "distr" ].
\end{coq_example}

\item
\begin{tabular}[t]{@{}l}
  {\tt SearchAbout {\term} inside {\module$_1$} \ldots{} {\module$_n$}.} \\
  {\tt SearchAbout [ \nelist{\textrm{\textsl{qualid-or-string}}}{} ]
    inside {\module$_1$} \ldots{} {\module$_n$}.}
\end{tabular}

This restricts the search to constructions defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\item
\begin{tabular}[t]{@{}l}
  {\tt SearchAbout {\term} outside {\module$_1$}...{\module$_n$}.} \\
  {\tt SearchAbout [ \nelist{\textrm{\textsl{qualid-or-string}}}{} ]
     outside {\module$_1$}...{\module$_n$}.}
\end{tabular}

This restricts the search to constructions not defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\end{Variants}

\subsection[\tt SearchPattern {\term}.]{\tt SearchPattern {\term}.\comindex{SearchPattern}}

This command displays the name and type of all theorems of the current
context whose statement's conclusion matches the expression {\term}
where holes in the latter are denoted by ``{\texttt \_}''.

\begin{coq_example}
Require Import Arith.
SearchPattern (_ + _ = _ + _).
\end{coq_example}

Patterns need not be linear: you can express that the same expression
must occur in two places by using pattern variables `{\texttt
?{\ident}}''.

\begin{coq_example}
Require Import Arith.
SearchPattern (?X1 + _ = _ + ?X1).
\end{coq_example}

\begin{Variants}
\item {\tt SearchPattern {\term} inside
{\module$_1$} \ldots{} {\module$_n$}.}\comindex{SearchPattern \ldots{} inside
\ldots{}}

This restricts the search to constructions defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\item {\tt SearchPattern {\term} outside {\module$_1$} \ldots{} {\module$_n$}.}\comindex{SearchPattern \ldots{} outside \ldots{}}

This restricts the search to constructions not defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\end{Variants}

\subsection[\tt SearchRewrite {\term}.]{\tt SearchRewrite {\term}.\comindex{SearchRewrite}}

This command displays the name and type of all theorems of the current
context whose statement's conclusion is an equality of which one side matches
the expression {\term =}. Holes in {\term} are denoted by ``{\texttt \_}''.

\begin{coq_example}
Require Import Arith.
SearchRewrite (_ + _ + _).
\end{coq_example}

\begin{Variants}
\item {\tt SearchRewrite {\term} inside
{\module$_1$} \ldots{} {\module$_n$}.}

This restricts the search to constructions defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\item {\tt SearchRewrite {\term} outside {\module$_1$} \ldots{} {\module$_n$}.}

This restricts the search to constructions not defined in modules
{\module$_1$} \ldots{} {\module$_n$}.

\end{Variants}

% \subsection[\tt SearchIsos {\term}.]{\tt SearchIsos {\term}.\comindex{SearchIsos}}
% \label{searchisos}
% \texttt{SearchIsos} searches terms by their type modulo isomorphism.
% This command displays the full name of all constants, variables,
% inductive types, and inductive constructors of the current
% context whose type is isomorphic to {\term} modulo the contextual part of the
% following axiomatization (the mutual inductive types with one constructor,
% without implicit arguments, and for which projections exist, are regarded as a
% sequence of $\sa{}$):


% \begin{tabbing}
% \ \ \ \ \=11.\ \=\kill
% \>1.\>$A=B\mx{ if }A\stackrel{\bt{}\io{}}{\lra{}}B$\\
% \>2.\>$\sa{}x:A.B=\sa{}y:A.B[x\la{}y]\mx{ if }y\not\in{}FV(\sa{}x:A.B)$\\
% \>3.\>$\Pi{}x:A.B=\Pi{}y:A.B[x\la{}y]\mx{ if }y\not\in{}FV(\Pi{}x:A.B)$\\
% \>4.\>$\sa{}x:A.B=\sa{}x:B.A\mx{ if }x\not\in{}FV(A,B)$\\
% \>5.\>$\sa{}x:(\sa{}y:A.B).C=\sa{}x:A.\sa{}y:B[y\la{}x].C[x\la{}(x,y)]$\\
% \>6.\>$\Pi{}x:(\sa{}y:A.B).C=\Pi{}x:A.\Pi{}y:B[y\la{}x].C[x\la{}(x,y)]$\\
% \>7.\>$\Pi{}x:A.\sa{}y:B.C=\sa{}y:(\Pi{}x:A.B).(\Pi{}x:A.C[y\la{}(y\sm{}x)]$\\
% \>8.\>$\sa{}x:A.unit=A$\\
% \>9.\>$\sa{}x:unit.A=A[x\la{}tt]$\\
% \>10.\>$\Pi{}x:A.unit=unit$\\
% \>11.\>$\Pi{}x:unit.A=A[x\la{}tt]$
% \end{tabbing}

% For more informations about the exact working of this command, see
% \cite{Del97}.

\subsection[\tt Locate {\qualid}.]{\tt Locate {\qualid}.\comindex{Locate}
\label{Locate}}
This command displays the full name of the qualified identifier {\qualid}
and consequently the \Coq\ module in which it is defined.

\begin{coq_eval}
(*************** The last line should produce **************************)
(*********** Error: I.Dont.Exist not a defined object ******************)
\end{coq_eval}
\begin{coq_eval}
Set Printing Depth 50.
\end{coq_eval}
\begin{coq_example}
Locate nat.
Locate Datatypes.O.
Locate Init.Datatypes.O.
Locate Coq.Init.Datatypes.O.
Locate I.Dont.Exist.
\end{coq_example}

\SeeAlso Section \ref{LocateSymbol}

\subsection{The {\sc Whelp} searching tool
\label{Whelp}}

{\sc Whelp} is an experimental searching and browsing tool for the
whole {\Coq} library and the whole set of {\Coq} user contributions.
{\sc Whelp} requires a browser to work. {\sc Whelp} has been developed
at the University of Bologna as part of the HELM\footnote{Hypertextual
Electronic Library of Mathematics} and MoWGLI\footnote{Mathematics on
the Web, Get it by Logics and Interfaces} projects.  It can be invoked
directly from the {\Coq} toplevel or from {\CoqIDE}, assuming a
graphical environment is also running. The browser to use can be
selected by setting the environment variable {\tt
COQREMOTEBROWSER}. If not explicitly set, it defaults to
\verb!firefox -remote "OpenURL(%s)"!  or
\verb!C:\\PROGRA~1\\INTERN~1\\IEXPLORE %s!, depending on the
underlying operating system (in the command, the string \verb!%s!
serves as metavariable for the url to open).
The Whelp tool relies on a dedicated Whelp server and on another server
called Getter that retrieves formal documents. The default Whelp server name
can be obtained using the command {\tt Test Whelp Server}
\comindex{Test Whelp Server} and the default Getter can be obtained
using the command: {\tt Test Whelp Getter} \comindex{Test Whelp
Getter}. The Whelp server name can be changed using the command:

\smallskip
\noindent {\tt Set Whelp Server {\str}}.\\
where {\str} is a URL (e.g. {\tt http://mowgli.cs.unibo.it:58080}).
\comindex{Set Whelp Server}
\smallskip

\noindent The Getter can be changed using the command:
\smallskip

\noindent {\tt Set Whelp Getter {\str}}.\\
where {\str} is a URL (e.g. {\tt http://mowgli.cs.unibo.it:58081}).  
\comindex{Set Whelp Getter}

\bigskip

The {\sc Whelp} commands are:

\subsubsection{\tt Whelp Locate "{\sl reg\_expr}".
\comindex{Whelp Locate}}

This command opens a browser window and displays the result of seeking
for all names that match the regular expression {\sl reg\_expr} in the
{\Coq} library and user contributions. The regular expression can
contain the special operators are * and ? that respectively stand for
an arbitrary substring and for exactly one character.

\variant {\tt Whelp Locate {\ident}.}\\
This is equivalent to {\tt Whelp Locate "{\ident}"}.

\subsubsection{\tt Whelp Match {\pattern}.
\comindex{Whelp Match}}

This command opens a browser window and displays the result of seeking
for all statements that match the pattern {\pattern}. Holes in the
pattern are represented by the wildcard character ``\_''.

\subsubsection[\tt Whelp Instance {\pattern}.]{\tt Whelp Instance {\pattern}.\comindex{Whelp Instance}}

This command opens a browser window and displays the result of seeking
for all statements that are instances of the pattern {\pattern}. The
pattern is here assumed to be an universally quantified expression.

\subsubsection[\tt Whelp Elim {\qualid}.]{\tt Whelp Elim {\qualid}.\comindex{Whelp Elim}}

This command opens a browser window and displays the result of seeking
for all statements that have the ``form'' of an elimination scheme
over the type denoted by {\qualid}.

\subsubsection[\tt Whelp Hint {\term}.]{\tt Whelp Hint {\term}.\comindex{Whelp Hint}}

This command opens a browser window and displays the result of seeking
for all statements that can be instantiated so that to prove the
statement {\term}.

\variant {\tt Whelp Hint.}\\ This is equivalent to {\tt Whelp Hint
{\sl goal}} where {\sl goal} is the current goal to prove. Notice that
{\Coq} does not send the local environment of definitions to the {\sc
Whelp} tool so that it only works on requests strictly based on, only,
definitions of the standard library and user contributions.

\section{Loading files}

\Coq\ offers the possibility of loading different
parts of a whole development stored in separate files. Their contents
will be loaded as if they were entered from the keyboard. This means
that the loaded files are ASCII files containing sequences of commands
for \Coq's toplevel. This kind of file is called a {\em script} for
\Coq\index{Script file}. The standard (and default) extension of
\Coq's script files is {\tt .v}.

\subsection[\tt Load {\ident}.]{\tt Load {\ident}.\comindex{Load}\label{Load}}
This command loads the file named {\ident}{\tt .v}, searching
successively in each of the directories specified in the {\em
  loadpath}. (see Section~\ref{loadpath})

\begin{Variants}
\item {\tt Load {\str}.}\label{Load-str}\\
  Loads the file denoted by the string {\str}, where {\str} is any
  complete filename. Then the \verb.~. and {\tt ..}
  abbreviations are allowed as well as shell variables. If no
  extension is specified, \Coq\ will use the default extension {\tt
    .v}
\item {\tt Load Verbose {\ident}.}, 
  {\tt Load Verbose {\str}}\\
  \comindex{Load Verbose}
  Display, while loading, the answers of \Coq\ to each command
  (including tactics) contained in the loaded file
  \SeeAlso Section~\ref{Begin-Silent}
\end{Variants}

\begin{ErrMsgs}
\item \errindex{Can't find file {\ident} on loadpath}
\end{ErrMsgs}

\section[Compiled files]{Compiled files\label{compiled}\index{Compiled files}}

This feature allows to build files for a quick loading. When loaded,
the commands contained in a compiled file will not be {\em replayed}.
In particular, proofs will not be replayed. This avoids a useless
waste of time.

\Rem A module containing an opened section cannot be compiled. 

% \subsection{\tt Compile Module {\ident}.}
% \index{Modules}
% \comindex{Compile Module}
% \index{.vo files}
% This command loads the file
% {\ident}{\tt .v} and plays the script it contains. Declarations,
% definitions and proofs it contains are {\em "packaged"} in a compiled
% form: the {\em module} named {\ident}.
% A file {\ident}{\tt .vo} is then created.
% The file {\ident}{\tt .v} is searched according to the
% current loadpath.
% The {\ident}{\tt .vo} is then written in the directory where
% {\ident}{\tt .v} was found.

% \begin{Variants}
% \item \texttt{Compile Module {\ident} {\str}.}\\ 
%   Uses the file {\str}{\tt .v} or {\str} if the previous one does not
%   exist to build the module {\ident}. In this case, {\str} is any
%   string giving a filename in the UNIX sense (see section
%   \ref{Load-str}). 
%   \Warning The given filename can not contain other caracters than
%   the caracters of \Coq's identifiers : letters or digits or the
%   underscore symbol ``\_''.

% \item \texttt{Compile Module Specification {\ident}.}\\
%   \comindex{Compile Module Specification}
%   Builds a specification module: only the types of terms are stored
%   in the module. The bodies (the proofs) are {\em not} written
%   in the module. In that case, the file created is {\ident}{\tt .vi}.
%   This is only useful when proof terms take too much place in memory
%   and are not necessary.
  
% \item \texttt{Compile Verbose Module {\ident}.}\\ 
%   \comindex{Compile Verbose Module}
%   Verbose version of Compile: shows the contents of the file being
%   compiled.
% \end{Variants}

% These different variants can be combined.


% \begin{ErrMsgs}
% \item \texttt{You cannot open a module when there are things other than}\\
%   \texttt{Modules and Imports in the context.}\\ 
%   The only commands allowed before a {Compile Module} command are {\tt
%     Require},\\
%   {\tt Read Module} and {\tt Import}. Actually, The normal way to
%   compile modules is by the {\tt coqc} command (see chapter
%   \ref{Addoc-coqc}).
% \end{ErrMsgs}

% \SeeAlso sections \ref{Opaque}, \ref{loadpath}, chapter
% \ref{Addoc-coqc}

%\subsection[\tt Import {\qualid}.]{\tt Import {\qualid}.\comindex{Import}}
%\label{Import}

%%%%%%%%%%%%
% Import and Export described in RefMan-mod.tex
% the minor difference (to avoid multiple Exporting of libraries) in
% the treatment of normal modules and libraries by Export omitted


\subsection[\tt Require {\dirpath}.]{\tt Require {\dirpath}.\label{Require}
\comindex{Require}}

This command looks in the loadpath for a file containing module
{\dirpath}, then loads and opens (imports) its contents.
More precisely, if {\dirpath} splits into a library dirpath {\dirpath'} and a module name {\textsl{ident}}, then the file {\ident}{\tt .vo} is searched in a physical path mapped to the logical path {\dirpath'}.

TODO: effect on the name table.

% The implementation file ({\ident}{\tt .vo}) is searched first,
% then the specification file ({\ident}{\tt .vi}) in case of failure.
If the module required has already been loaded, \Coq\ 
simply opens it (as {\tt Import {\dirpath}} would do it).
%If the module required is already loaded and open, \Coq\ 
%displays the following warning: {\tt {\ident} already imported}.

If a module {\it A} contains a command {\tt Require} {\it B} then the
command {\tt Require} {\it A} loads the module {\it B} but does not
open it (See the {\tt Require Export} variant below).

\begin{Variants}
\item {\tt Require Export {\qualid}.}\\
  \comindex{Require Export}
  This command acts as {\tt Require} {\qualid}.  But if a module {\it
    A} contains a command {\tt Require Export} {\it B}, then the
  command {\tt Require} {\it A} opens the module {\it B} as if the
  user would have typed {\tt Require}{\it B}.
% \item {\tt Require $[$ Implementation $|$ Specification $]$ {\qualid}.}\\
%   \comindex{Require Implementation}
%   \comindex{Require Specification}
%   Is the same as {\tt Require}, but specifying explicitly the
%   implementation ({\tt.vo} file) or the specification ({\tt.vi}
%   file).

% Redundant ?
% \item {\tt Require {\qualid} {\str}.}\\ 
%   Specifies the file to load as being {\str} but containing module
%  {\qualid}. 
% The opened module is still {\ident} and therefore must have been loaded.
\item {\tt Require {\qualid} {\str}.}\\ 
  Specifies the file to load as being {\str} but containing module
  {\qualid} which is then opened.
\end{Variants}

These different variants can be combined.

\begin{ErrMsgs}

\item \errindex{Cannot load {\ident}: no physical path bound to {\dirpath}}

\item \errindex{Can't find module toto on loadpath}

  The command did not find the file {\tt toto.vo}. Either {\tt
    toto.v} exists but is not compiled or {\tt toto.vo} is in a directory
  which is not in your {\tt LoadPath} (see Section~\ref{loadpath}).

\item \errindex{Bad magic number}

  \index{Bad-magic-number@{\tt Bad Magic Number}}
  The file {\tt{\ident}.vo} was found but either it is not a \Coq\
  compiled module, or it was compiled with an older and incompatible
  version of \Coq.
\end{ErrMsgs}

\SeeAlso Chapter~\ref{Addoc-coqc}

\subsection[\tt Print Modules.]{\tt Print Modules.\comindex{Print Modules}}
This command shows the currently loaded and currently opened
(imported) modules.

\subsection[\tt Declare ML Module {\str$_1$} .. {\str$_n$}.]{\tt Declare ML Module {\str$_1$} .. {\str$_n$}.\comindex{Declare ML Module}}
This commands loads the Objective Caml compiled files {\str$_1$} \dots
{\str$_n$} (dynamic link). It is mainly used to load tactics
dynamically.
% (see Chapter~\ref{WritingTactics}).
 The files are
searched into the current Objective Caml loadpath (see the command {\tt
Add ML Path} in the Section~\ref{loadpath}).  Loading of Objective Caml
files is only possible under the bytecode version of {\tt coqtop}
(i.e. {\tt coqtop} called with options {\tt -byte}, see chapter 
\ref{Addoc-coqc}).

\begin{ErrMsgs}
\item \errindex{File not found on loadpath : \str}
\item \errindex{Loading of ML object file forbidden in a native Coq}
\end{ErrMsgs}

\subsection[\tt Print ML Modules.]{\tt Print ML Modules.\comindex{Print ML Modules}}
This print the name of all \ocaml{} modules loaded with \texttt{Declare
  ML Module}. To know from where these module were loaded, the user
should use the command \texttt{Locate File} (see Section~\ref{Locate File})

\section[Loadpath]{Loadpath\label{loadpath}\index{Loadpath}}

There are currently two loadpaths in \Coq. A loadpath where seeking
{\Coq} files (extensions {\tt .v} or {\tt .vo} or {\tt .vi}) and one where
seeking Objective Caml files. The default loadpath contains the
directory ``\texttt{.}'' denoting the current directory and mapped to the empty logical path (see Section~\ref{LongNames}).

\subsection[\tt Pwd.]{\tt Pwd.\comindex{Pwd}\label{Pwd}}
This command displays the current working directory.

\subsection[\tt Cd {\str}.]{\tt Cd {\str}.\comindex{Cd}}
This command changes the current directory according to {\str} 
which can be any valid path.

\begin{Variants}
\item {\tt Cd.}\\
  Is equivalent to {\tt Pwd.}
\end{Variants}

\subsection[\tt Add LoadPath {\str} as {\dirpath}.]{\tt Add LoadPath {\str} as {\dirpath}.\comindex{Add LoadPath}\label{AddLoadPath}}

This command adds the path {\str} to the current {\Coq} loadpath and
maps it to the logical directory {\dirpath}, which means that every
file {\tt M.v} physically lying in directory {\str} becomes accessible
through logical name ``{\dirpath}{\tt{.M}}''. 

\Rem {\tt Add LoadPath} also adds {\str} to the current ML loadpath.

\begin{Variants}
\item {\tt Add LoadPath {\str}.}\\
Performs as {\tt Add LoadPath {\str} as {\dirpath}} but for the empty directory path.
\end{Variants}

\subsection[\tt Add Rec LoadPath {\str} as {\dirpath}.]{\tt Add Rec LoadPath {\str} as {\dirpath}.\comindex{Add Rec LoadPath}\label{AddRecLoadPath}}
This command adds the directory {\str} and all its subdirectories 
to the current \Coq\ loadpath. The top directory {\str} is mapped to the logical directory {\dirpath} while any subdirectory {\textsl{pdir}} is mapped to logical directory {\dirpath}{\tt{.pdir}} and so on.

\Rem {\tt Add Rec LoadPath} also recursively adds {\str} to the current ML loadpath.

\begin{Variants}
\item {\tt Add Rec LoadPath {\str}.}\\
Works as {\tt Add Rec LoadPath {\str} as {\dirpath}} but for the empty logical directory path.
\end{Variants}

\subsection[\tt Remove LoadPath {\str}.]{\tt Remove LoadPath {\str}.\comindex{Remove LoadPath}}
This command removes the path {\str} from the current \Coq\ loadpath.

\subsection[\tt Print LoadPath.]{\tt Print LoadPath.\comindex{Print LoadPath}}
This command displays the current \Coq\ loadpath.

\subsection[\tt Add ML Path {\str}.]{\tt Add ML Path {\str}.\comindex{Add ML Path}}
This command adds the path {\str} to the current Objective Caml loadpath (see
the command {\tt Declare ML Module} in the Section~\ref{compiled}).

\Rem This command is implied by {\tt Add LoadPath {\str} as {\dirpath}}.

\subsection[\tt Add Rec ML Path {\str}.]{\tt Add Rec ML Path {\str}.\comindex{Add Rec ML Path}}
This command adds the directory {\str} and all its subdirectories 
to the current Objective Caml loadpath (see
the command {\tt Declare ML Module} in the Section~\ref{compiled}).

\Rem This command is implied by {\tt Add Rec LoadPath {\str} as {\dirpath}}.

\subsection[\tt Print ML Path {\str}.]{\tt Print ML Path {\str}.\comindex{Print ML Path}}
This command displays the current Objective Caml loadpath.
This command makes sense only under the bytecode version of {\tt
coqtop}, i.e. using option {\tt -byte} (see the
command {\tt Declare ML Module} in the section
\ref{compiled}).

\subsection[\tt Locate File {\str}.]{\tt Locate File {\str}.\comindex{Locate
  File}\label{Locate File}}
This command displays the location of file {\str} in the current loadpath.
Typically, {\str} is a \texttt{.cmo} or \texttt{.vo} or \texttt{.v} file.

\subsection[\tt Locate Library {\dirpath}.]{\tt Locate Library {\dirpath}.\comindex{Locate Library}}
This command gives the status of the \Coq\ module {\dirpath}. It tells if the
module is loaded and if not searches in the load path for a module
of logical name {\dirpath}.

\section{States and Reset}

\subsection[\tt Reset \ident.]{\tt Reset \ident.\comindex{Reset}}
This command removes all the objects in the environment since \ident\ 
was introduced, including \ident. \ident\ may be the name of a defined
or declared object as well as the name of a section.  One cannot reset
over the name of a module or of an object inside a module.

\begin{ErrMsgs}
\item \ident: \errindex{no such entry}
\end{ErrMsgs}

\subsection[\tt Back.]{\tt Back.\comindex{Back}}

This commands undoes all the effects of the last vernacular
command. This does not include commands that only access to the
environment like those described in the previous sections of this
chapter (for instance {\tt Require} and {\tt Load} can be undone, but
not {\tt Check} and {\tt Locate}). Commands read from a vernacular
file are considered as a single command.

\begin{Variants}
\item {\tt Back $n$} \\
  Undoes $n$ vernacular commands.
\end{Variants}

\begin{ErrMsgs}
\item \errindex{Reached begin of command history} \\
  Happens when there is vernacular command to undo.
\end{ErrMsgs}

\subsection[\tt Backtrack $\num_1$ $\num_2$ $\num_3$.]{\tt Backtrack $\num_1$ $\num_2$ $\num_3$.\comindex{Backtrack}}

This command is dedicated for the use in graphical interfaces.  It
allows to backtrack to a particular \emph{global} state, i.e.
typically a state corresponding to a previous line in a script. A
global state includes declaration environment but also proof
environment (see Chapter~\ref{Proof-handling}). The three numbers
$\num_1$, $\num_2$ and $\num_3$ represent the following:
\begin{itemize}
\item $\num_3$: Number of \texttt{Abort} to perform, i.e. the number
  of currently opened nested proofs that must be cancelled (see
  Chapter~\ref{Proof-handling}).
\item $\num_2$: \emph{Proof state number} to unbury once aborts have
  been done. Coq will compute the number of \texttt{Undo} to perform
  (see Chapter~\ref{Proof-handling}).
\item $\num_1$: Environment state number to unbury, Coq will compute
  the number of \texttt{Back} to perform.
\end{itemize}


\subsubsection{How to get state numbers?}
\label{sec:statenums}


Notice that when in \texttt{-emacs} mode, \Coq\ displays the current
proof and environment state numbers in the prompt. More precisely the
prompt in \texttt{-emacs} mode is the following:

\verb!<prompt>! \emph{$id_i$} \verb!<! $\num_1$
\verb!|! $id_1$\verb!|!$id_2$\verb!|!\dots\verb!|!$id_n$
\verb!|! $\num_2$ \verb!< </prompt>!

Where:

\begin{itemize}
\item \emph{$id_i$} is the name of the current proof (if there is
  one, otherwise \texttt{Coq} is displayed, see
Chapter~\ref{Proof-handling}).
\item $\num_1$ is the environment state number after the last
  command.
\item $\num_2$ is the proof state number after the last
  command.
\item $id_1$ $id_2$ \dots $id_n$ are the currently opened proof names
  (order not significant).
\end{itemize}

It is then possible to compute the \texttt{Backtrack} command to
unbury the state corresponding to a particular prompt. For example,
suppose the current prompt is:

\verb!<! goal4 \verb!<! 35
\verb!|!goal1\verb!|!goal4\verb!|!goal3\verb!|!goal2\verb!|! 
\verb!|!8 \verb!< </prompt>!

and we want to backtrack to a state labelled by:

\verb!<! goal2 \verb!<! 32
\verb!|!goal1\verb!|!goal2
\verb!|!12 \verb!< </prompt>!

We have to perform \verb!Backtrack 32 12 2! , i.e. perform 2
\texttt{Abort}s (to cancel goal4 and goal3), then rewind proof until
state 12 and finally go back to environment state 32. Notice that this
supposes that proofs are nested in a regular way (no \texttt{Resume} or
\texttt{Suspend} commands).

\begin{Variants}
\item {\tt BackTo n}. \comindex{BackTo}\\
  Is a more basic form of \texttt{Backtrack} where only the first
  argument (global environment number) is given, no \texttt{abort} and
  no \texttt{Undo} is performed.
\end{Variants}

\subsection[\tt Restore State \str.]{\tt Restore State \str.\comindex{Restore State}}
  Restores the state contained in the file \str.

\begin{Variants}
\item {\tt Restore State \ident}\\
 Equivalent to {\tt Restore State "}{\ident}{\tt .coq"}.
\item {\tt Reset Initial.}\comindex{Reset Initial}\\ 
  Goes back to the initial state (like after the command {\tt coqtop},
  when the interactive session began). This command is only available
  interactively.
\end{Variants}

\subsection[\tt Write State \str.]{\tt Write State \str.\comindex{Write State}}
Writes the current state into a file \str{} for
use in a further session. This file can be given as the {\tt
  inputstate} argument of the commands {\tt coqtop} and {\tt coqc}.

\begin{Variants}
\item {\tt Write State \ident}\\
 Equivalent to {\tt Write State "}{\ident}{\tt .coq"}.
 The state is saved in the current directory (see Section~\ref{Pwd}).
\end{Variants}

\section{Quitting and debugging}

\subsection[\tt Quit.]{\tt Quit.\comindex{Quit}}
This command permits to quit \Coq.

\subsection[\tt Drop.]{\tt Drop.\comindex{Drop}\label{Drop}}

This is used mostly as a debug facility by \Coq's implementors
and does not concern the casual user.
This command permits to leave {\Coq} temporarily and enter the
Objective Caml toplevel. The Objective Caml command:

\begin{flushleft}
\begin{verbatim}
#use "include";;
\end{verbatim}
\end{flushleft}

\noindent add the right loadpaths and loads some toplevel printers for
all abstract types of \Coq - section\_path, identfifiers, terms, judgements,
\dots. You can also use the file \texttt{base\_include} instead,
that loads only the pretty-printers for section\_paths and
identifiers.
% See Section~\ref{test-and-debug} more information on the
% usage of the toplevel.
You can return back to \Coq{} with the command: 

\begin{flushleft}
\begin{verbatim}
go();;
\end{verbatim}
\end{flushleft}

\begin{Warnings}
\item It only works with the bytecode version of {\Coq} (i.e. {\tt coqtop} called with option {\tt -byte}, see the contents of Section~\ref{binary-images}).
\item You must have compiled {\Coq} from the source package and set the
  environment variable \texttt{COQTOP} to the root of your copy of the sources (see Section~\ref{EnvVariables}).
\end{Warnings}

\subsection[\tt Time \textrm{\textsl{command}}.]{\tt Time \textrm{\textsl{command}}.\comindex{Time}
\label{time}}
This command executes the vernac command \textrm{\textsl{command}}
and display the time needed to execute it.

\section{Controlling display}

\subsection[\tt Set Silent.]{\tt Set Silent.\comindex{Set Silent}
\label{Begin-Silent}
\index{Silent mode}}
This command turns off the normal displaying.

\subsection[\tt Unset Silent.]{\tt Unset Silent.\comindex{Unset Silent}}
This command turns the normal display on.

\subsection[\tt Set Printing Width {\integer}.]{\tt Set Printing Width {\integer}.\comindex{Set Printing Width}}
This command sets which left-aligned part of the width of the screen
is used for display. 

\subsection[\tt Unset Printing Width.]{\tt Unset Printing Width.\comindex{Unset Printing Width}}
This command resets the width of the screen used for display to its
default value (which is 78 at the time of writing this documentation).

\subsection[\tt Test Printing Width.]{\tt Test Printing Width.\comindex{Test Printing Width}}
This command displays the current screen width used for display.

\subsection[\tt Set Printing Depth {\integer}.]{\tt Set Printing Depth {\integer}.\comindex{Set Printing Depth}}
This command sets the nesting depth of the formatter used for
pretty-printing. Beyond this depth, display of subterms is replaced by
dots. 

\subsection[\tt Unset Printing Depth.]{\tt Unset Printing Depth.\comindex{Unset Printing Depth}}
This command resets the nesting depth of the formatter used for
pretty-printing to its default value (at the
time of writing this documentation, the default value is 50).

\subsection[\tt Test Printing Depth.]{\tt Test Printing Depth.\comindex{Test Printing Depth}}
This command displays the current nesting depth used for display.

%\subsection{\tt Explain ...}
%Not yet documented.

%\subsection{\tt Go ...}
%Not yet documented.

%\subsection{\tt Abstraction ...}
%Not yet documented.

\section{Controlling the conversion algorithm}

{\Coq} comes with two algorithms to check the convertibility of types
(see Section~\ref{convertibility}). The first algorithm lazily
compares applicative terms while the other is a brute-force but efficient
algorithm that first normalizes the terms before comparing them.  The
second algorithm is based on a bytecode representation of terms
similar to the bytecode representation used in the ZINC virtual
machine~\cite{Leroy90}. It is specially useful for intensive
computation of algebraic values, such as numbers, and for reflexion-based
tactics.

\subsection{\tt Set Virtual Machine
\label{SetVirtualMachine}
\comindex{Set Virtual Machine}}

This activates the bytecode-based conversion algorithm.

\subsection{\tt Unset Virtual Machine
\comindex{Unset Virtual Machine}}

This deactivates the bytecode-based conversion algorithm.

\subsection{\tt Test Virtual Machine
\comindex{Test Virtual Machine}}

This tells if the bytecode-based conversion algorithm is
activated. The default behavior is to have the bytecode-based
conversion algorithm deactivated.

\SeeAlso sections~\ref{vmcompute} and~\ref{vmoption}.

% $Id$ 

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Reference-Manual"
%%% End: 
